### 2025-04-22 paper/main.tex
change affilation

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{authblk}

\title{Three Algorithms for Merging Navigable Graphs}
\author[1]{Alexander Ponomarenko}
\date{}

% \affil[2]{YugaByte LLC}
\affil[1]{HSE Univerity}

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\DeclareMathOperator*{\argminA}{arg\,min} % Jan Hlavacek
\DeclareMathOperator*{\argminB}{argmin}   % Jan Hlavacek
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{nameref}
\begin{document}

\maketitle

\section{Introduction}
In recent years, navigable small-world graphs have become a cornerstone of efficient approximate nearest neighbor (ANN) search in high-dimensional spaces, a problem prevalent in fields such as recommendation systems, image retrieval, and natural language processing. Algorithms like Hierarchical Navigable Small World (HNSW) \cite{hnsw} and its predecessor Navigable Small World (NSW) \cite{nsw2011,nsw2012,nsw2014} provide a robust and scalable framework for organizing data into graph structures that support fast similarity searches. However, one limitation inherent to these methods is the lack of a natural mechanism for node deletion, making the need for merging and compacting graphs crucial for maintaining the efficiency and accuracy of the search index over time.

The HNSW algorithm, widely adopted for its performance and scalability, supports dynamic insertions but lacks native support for node deletion. As databases grow and evolve, some nodes become outdated or irrelevant, requiring a process to filter out these deleted vertices. Merging operations allow for a form of "data compaction," where the graphs from two or more databases are merged, and nodes marked for deletion are omitted from the final structure. This process not only improves space efficiency but also enhances the accuracy of future searches by eliminating superfluous data points.

In this work, we propose three novel algorithms designed to efficiently merge navigable graphs, focusing on HNSW but extending applicability to other graph-based data structures like NSW \cite{nsw2011}, NSG \cite{NSG}. The merging algorithms—"Naive Merge," "Merge1," and "Merge2"—offer progressively advanced methods to combine graphs while preserving the properties that make navigable graphs effective for nearest neighbor search. These methods ensure that the resulting merged index excludes deleted nodes and maintains the structural integrity needed for fast search operations.

\section {Search}

We start from the basic search over the sigle graph.

\begin{algorithm}
\caption{\textsc{LocalSearch}($G, q, C, k, L$)}\label{alg:local_search}
\textbf{Require:} Graph $G = (V, E)$, query $q \in \mathbb{R}^d$, initial set of candidate vertexes $C \subset V$,  $k \in \mathbb{N}$, $L \in \mathbb{N}$ \\
\textbf{Ensure:} approximate k-nearest neighbors $V^* \subset V$
\begin{algorithmic}[1]
% \State $C \gets \textsc{InitializeCandidates}()$
\While{True}
    \State $u \gets$ nearest unvisited point to $q$ in $C$
    \State $U \gets \{v \mid (u, v) \in E\}$
    \For{$v \in U$}
        \If{$v$ is not visited}
            \State $C \gets C \cup \{v\}$
        \EndIf
    \EndFor
    \If{$|C| > L$}
        \State $C \gets$ top $L$ nearest points to $q$ in $C$
    \EndIf
    \If{$C$ is not updated}
        \State \textbf{break}
    \EndIf
\EndWhile
\State \Return nearest point to $q$ in $C$
\end{algorithmic}
\end{algorithm}

$L$ is a parameter that controls the width of search. 
It helps to avoid local minima vertexes. In [HNSW] paper it is named \textbf{ef} parameter (expansion factor).

\begin{algorithm}
\caption{\textsc{HNSW-Search}($\mathcal{H}, q, v_0, k, L, \ell $)}\label{alg:hnsw_search}
\textbf{Require:} The HNSW layers represented as  a sequence of graphs $\mathcal{H} = (G_i)_{i=0}^{l_{max}} )$, query $q \in \mathbb{R}^d$, initial vertex $v_0 \in V$,  $k \in \mathbb{N}$, $L \in \mathbb{N}$, the layer number in which the search should be performed $\ell$ \\    
\textbf{Ensure:} approximate k-nearest neighbors $V^* \subset V$
\begin{algorithmic}[1]
% \State $C \gets \textsc{InitializeCandidates}()$

\State $v^* \gets v_0$ 
\For{$i = l_{max} \; \textbf{to} \; \ell $}
    \State $v^* \gets \textsc{LocalSearch}(G=G_i, q=q, C={v^*}, k=1, L=L)$
\EndFor

\State \Return \textsc{LocalSearch}($ G=G_{\ell}, q=q, C={v^*}, k=k, L=L$)
\end{algorithmic}
\end{algorithm}

\section {Neighborhood Construction Strategies}

There are different strategies to form a vertex neighborhood. Our proposed merge algorithms are independent in neighborhood procedures. Merge algorithm can use a neighborhood construction procedure as an input parameter.

\begin{algorithm}
\caption{\textsc{KNN-Neighborhood-Construction}$(v^*,C, k)$}
\label{alg:knntrategy}
\textbf{Require:} vertex $v^* \in V$, neighbor candidates $C \subseteq V$, $k \in \mathbb{N} $\\
\textbf{Ensure:} selected neighbors $C' \subseteq C$
\begin{algorithmic}[1]

    \State $C' \gets \text{k-closest neighbors of } C \text{ to } v^*$
    \State \Return $C'$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{\textsc{RNG-Neighborhood-Construction}$(v^*, C, m)$}
\label{alg:rngstrategy}
\textbf{Require:} vertex $v^* \in V$, neighbor candidates $C \subseteq V$, maximum neighborhood size $m \in \mathbb{N}$\\
\textbf{Ensure:} selected neighbors $C' \subseteq C$
\begin{algorithmic}[1]
    \State sort $v \in C$ in ascending order of $\rho(v^*,v)$
    \State $C' \gets \emptyset$
    \For{$v \in U$}
        \State $f \gets \text{true}$
        \For{$w \in C'$}
            \If{$\delta(v^*,v) \geq \delta(v,w)$}
                \State $f \gets \text{false}$
                \State \textbf{break}
            \EndIf
        \EndFor
        \If{$f$}
            \State $C' \gets C' \cup \{v\}$
        \EndIf

        \If{$|C'| \geq m $}
            \State \textbf{break}
        \EndIf
    \EndFor
    \State \Return $C'$
\end{algorithmic}
\end{algorithm}

\section{Merge Algorithms}

We provide 3 merge algorithms: "Naive Merge", "Merge1", and "Merge2". We also refer to these algorithm as layer-merge algorithms. "Naive Merge" is a straight forward method, while "Merge1" and "Merge2" are more advanced. Together, these two algorithms constitute the main contribution of the work. The HNSW structure is a sequence of graphs, called layers. We start from description of a general algorithm for merging multi layers structure.

\subsection{HNSW General Merge}

HNSW-General-Merge algorithm \ref{alg:general_merge} is a straight forward. It uses one of the layer-merge algorithm to merge each HNSW layer. To able layer-merge algorithm to perform fast search using multi-layer structure of HNSW, HNSW-General-Merge algorithm passes to layer merge algorithm both HNSW structures $\mathcal{H}_a, \mathcal{H}_b$, and the layer number $\ell=i$ to which the merge should be performed. An order of merging layers doesn't mater. Thus merging of layers can be done in parallel.

\begin{algorithm}
\caption{\textsc{HNSW-General-Merge}($\mathcal{H}_a, \mathcal{H}_b$)}\label{alg:general_merge}
\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$ \\
\textbf{Ensure:}  Merged HNSW graphs $\mathcal{H}_c = (G^c_i)$ 
\begin{algorithmic}[1]

\For{$i = 0 \textbf{ to } \max(|\mathcal{H}_a|, |\mathcal{H}_b|) $}
    \State $G^c_i \gets \text{Merge}(\mathcal{H}_a, \mathcal{H}_b, \ell=i, \text{algorithm specific params})$
\EndFor

\State $\mathcal{H}_c \gets (G^c_i)_i^{\max(|\mathcal{H}_a|, |\mathcal{H}_b|)}$

\State \Return $\mathcal{H}^c$
\end{algorithmic}
\end{algorithm}

\subsection{Naive Merge}
The Naive Merge algorithm \ref{alg:merge_naive} surprisingly is very simple. It takes one vertex $v^*$ by one from the graph $G^a$. Then algorithm prepares the candidate set $\mathcal{C}$ for a new neighborhood (lines 6,7). It uses \textsc{HNSW-Search} to get $m$-closest vertices to $v^*$ in the graph $G^b$, and combines with an old neighborhood of the vertex $v^*$. After that in the line 8 the candidate set $\mathcal{C}$ is passed into the neighborhood construction function, which selects the vertices for the new merged neighborhood for $v^*$.
Lines 9-12 contains the same logic for the forming new neighborhoods of graph $G^b$

\begin{algorithm}
\caption{\textsc{Merge-Naive}($\mathcal{H}_a, \mathcal{H}_b,  \ell, \text{search\_ef}$)}\label{alg:merge_naive}
\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i)$, $\mathcal{H}_b = (G^b_i)$, and the level number $\ell$, search parameter $\text{search\_ef}$ \\
\textbf{Ensure:} Merged graph $G^c$ 
\begin{algorithmic}[1]
\State $V^c \gets V^a \cup V^b$
\State $E^c \gets \emptyset$
\State $G^a \gets G^a_{\ell}$
\State $G^b \gets G^b_{\ell}$
% \State $m \gets \mathcal{H}_a.m_0$ if $\ell = 0$ else $\mathcal{H}_a.m$

\For{$v^* \in V^a$}
    \State $\mathcal{C}^b \gets \textsc{HNSW-Search}(\mathcal{H} = \mathcal{H}^b, q = v^*, v_0, k = m, L = \text{search\_ef}, \ell)$
    \State $\mathcal{C} \gets \{v : (v^*, v) \in E^a \} \cup \mathcal{C}^b$
    \State $E^c \gets E^c \cup \{(v^*, v) : v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
\EndFor

\For{$v^* \in V^b$}
    \State $\mathcal{C}^a \gets \textsc{HNSW-Search}(\mathcal{H} = \mathcal{H}^a, q = v^*, v_0, k = m, L = \text{search\_ef}, \ell)$
    \State $\mathcal{C} \gets \{v : (v^*, v) \in E^b \} \cup \mathcal{C}^a$
    \State $E^c \gets E^c \cup \{(v^*, v) : v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
\EndFor

\State \Return $G^c = (V^c, E^c)$
\end{algorithmic}
\end{algorithm}

\subsection{Merge1}
The most effort of \textsc{Merge-Naive} algorithm lies in the set of neighborhood candidates from the opposite graph utilising \textsc{HNSW-Search} procedure, that traverses layers graph from the top-level down to layer number $\ell$ every time. The number of computations can be reduced if we select the next vertex $v^*$ close to the previous one (line 15), instead of choosing it randomly. Thus, for new $v^*$ the neighborhood candidates will be  also close to the previous candidates set. To search this new neighborhood candidates we can use \textsc{LocalSearch} procedure, that traversing the same graph staring from the previous neighborhood candidates set $\mathcal{P}^b$ (line 11,14). In the line 14 in set $\mathcal{P}^b$ we keep only $M$-closest to $v^*$ candidates.\\
In the line 15 almost at each iteration we try select new $v^*$ vertex close to the previous  $v^*$ that was not already processed.  
To ensure that new $v^*$ vertex are not very far from the previous $v^*$ we bound the size of results of \textsc{LocalSearch} procedure controlling by "next\_step\_k" parameter. Once \textsc{LocalSearch} can't find enough close not processed vertex, in line 7 we choice new $v^*$ from the set $\mathcal{V}_{not\_done}$ randomly. \\
After we have processed all vertices from the the graph $G^a$. We do the same for vertices of graph $G^b$ (lines 21-22).

\begin{algorithm}
\caption{\textsc{Merge1}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge1}
\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$, the merging layer number $\ell$, the size of the forming neighborhoods $m \in \mathbb{N}$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$ \\
\textbf{Ensure:}  Merged graph $G^c$ 
\begin{algorithmic}[1]
\State $V^c \gets V^a \cup V^b$
\State $E^c \gets \emptyset$
\State $G^a \gets G^a_{\ell} $
\State $G^b \gets G^b_{\ell} $
\State $\mathcal{V}_{not\_done} \gets V^a$

\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State $v^* \gets \text{random choice from } \mathcal{V}_{not\_done}$

    \State $\mathcal{P}^b  \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^b, q=v^*, v_0, k=M, L=\text{jump\_ef}, \ell)$

    \While{True}
        \State $\mathcal{V}_{not\_done} \gets \mathcal{V}_{not\_done} \setminus \{v^*\}$

        \State $\mathcal{C}^b  \gets \textsc{LocalSearch}(G=G^b, q=v^*, C=\mathcal{P}^b , k=m, L=\text{local\_ef})$

        \State $\mathcal{C} \gets  \{v : (v^*, v) \in E^a \} \cup \mathcal{C}^b$

        \State $E^c \gets E^c \cup  \{ (v^*, v)  : v \in \textsc{NeighborhoodConstruction}(\mathcal{C}, v^*, m) \}$

        \State $\mathcal{P}^b \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_M \} $

        \State $\mathcal{C}^a  \gets \textsc{LocalSearch}(G=G^a, q=v^*, C=\{v^*\} , k=\text{next\_step\_k}, L=\text{next\_step\_ef})$

        \State $\mathcal{C}^a \gets \mathcal{C}^a \cap \mathcal{V}_{not\_done}$

        \If{$\mathcal{C}^a = \emptyset$}
            \State \textbf{break}
        \EndIf

        \State $v^* \gets \mathcal{C}^a_1$
    \EndWhile
\EndWhile

\State $\mathcal{V}_{not\_done} \gets V^b$

\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State Repeat the same process for $V^b$ with the roles of $\mathcal{H}_a$ and $\mathcal{H}_b$ swapped.
\EndWhile

\State \Return $G^c=(V^c,E^c)$

\end{algorithmic}
\end{algorithm}

\subsection{Merge2}

\textsc{Merge2} algorithm is similar to \textsc{Merge1} utilise \textsc{LocalSearch} procedure to reduce computation effort. The difference is that \textsc{Merge1} algorithm the next $v^*$ vertex choice from the same graph, while \textsc{Merge2} looks for new vertex $v^*$ in the both graphs $G^a$, and $G^b$. Thus in line 24  $v^*$  is chosen from the set $\mathcal{C}_{not\_done}$, which of not processed vertex from the both graph (line 21). The intuition laying behind \textsc{Merge2} algorithm is that allowing to choice new $v^*$ from the both graphs, we reduce the number of times what $v^*$ is chosen randomly, thereby minimise the number of using more expensive search procedure \textsc{HNSW-Search}.

\begin{algorithm}
\caption{\textsc{Merge2}($\mathcal{H}^a, \mathcal{H}^b,  \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge2}
\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$, the number of merged layers, the size of the forming neighborhoods $m \in \mathbb{N} $, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M \in \mathbb{N}$ \\
\textbf{Ensure:}  Merged graph $G^c$ 
\begin{algorithmic}[1]

\State $V^c \gets V^a \cup V^b$
\State $E^c \gets \emptyset$
\State $G^a \gets G^a_{\ell} $
\State $G^b \gets G^b_{\ell} $
% \State $m \gets \mathcal{H}_a.m_0 \text{ if } \ell = 0 \text{ else } \mathcal{H}_a.m$
\State $\mathcal{V}_{not\_done} \gets V^a \cup V^b$

\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State $v^* \gets \text{randomly choice from } \mathcal{V}_{not\_done}$

    \State $\mathcal{P}^a  \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^a, q=v^*, v_0, k, L=\text{jump\_ef}, \ell $)

    \State $\mathcal{P}^b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^b, q=v^*, v_0, k, L=\text{jump\_ef}, \ell $)

    \While{True}
        \State $\mathcal{V}_{not\_done} \gets \mathcal{V}_{not\_done} \setminus \{v^*\}$

        \State $ \mathcal{C}^a  \gets \textsc{LocalSearch}(G=G^a, q=v^*, C=\mathcal{P}^a  , k=m, L=\text{local\_ef})$

        \State $\mathcal{C}^b \gets \textsc{LocalSearch}(G=G^b, q=v^*, C=\mathcal{P}^b , k=m, L=\text{local\_ef})$

        \If{$v^* \in V^a $}
            \State $\mathcal{C}\gets  \{v : (v^*, v) \in E^a \} \cup  \mathcal{C}^b\}$
        \Else
            \State $\mathcal{C} \gets  \{v : (v^*, v) \in E^b \} \cup  \mathcal{C}^a\}$
        \EndIf

        \State $E^c \gets E^c \cup \{ (v^*, v) : v \in \text{neighborhood\_construction}(\mathcal{C}, v^*, m) \}$

        \State $\mathcal{C}^a_{not\_done} \gets \{\mathcal{C}^a_1, \mathcal{C}^a_2, ..., \mathcal{C}^a_{ \text{next\_step\_k} } \} \cap \mathcal{V}_{not\_done}$

        \State $\mathcal{C}^b_{not\_done} \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_{ \text{next\_step\_k} } \}  \cap \mathcal{V}_{not\_done}$

        \State $\mathcal{C}_{not\_done} \gets \mathcal{C}^a_{not\_done} \cup \mathcal{C}^b_{not\_done}$

        \If{$\mathcal{C}_{not\_done} = \emptyset$}
            \State \textbf{break}
        \EndIf

        \State $v^* \gets \underset{v \in \mathcal{C}_{not\_done}}{\mathrm{argmin}} \rho(v, v^*)$

        \State $\mathcal{P}_a \gets \mathcal{C}^a$
        \State $\mathcal{P}_b \gets \mathcal{C}^b$
    \EndWhile
\EndWhile

\State \Return $G^c=(V^c,E^c)$

\end{algorithmic}
\end{algorithm}

\section{Computational Experiments}
We implemented proposed algorithms using python. For experiments we use Sift1m dataset divided into two subset of 500k vectors at each. Our algorithm merge that two subsets of 500k vectors. In order to be independent of the implementation details and low level optimisation, we do comparison of the algorithms based on the number of distance computation during the merge process. \ref{fig:sfig1} shows that \textsc{Merge1} about 5 times less computations than naive algorithm. Surprisingly \textsc{Merge2} algorithm do slightly more computations than \textsc{Merge1}, nevertheless it significantly outperform \textsc{merge-naive} algorithm.\\
The other important characteristics of merging process is how accurate the merge is done. It can be easily verified for the case of \textsc{KNN-Neighborhood-Construction} procedure, by comparing the set of merged neighborhoods with k-nearest-neighbor set for each vertex. However to follow the same way in the case of \textsc{RNG-Neighborhood-Construction} is not good idea, because the \textsc{RNG-Neighborhood-Construction} is one of the possible heuristic for Voronoi-cell approximation. Therefore we measure the recall-distance computation trade off of \textsc{HNSW-Search} procedure on the HNSW graphs before merging, and after. As can be seen from the \ref{fig:sfig2} the recall/computational trade off of naive algorithm is slightly better than  \textsc{Merge1}, and  \textsc{Merge2} has.

\begin{figure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.\linewidth]{figs/merge-effort-comparison.pdf}
  \caption{Merging efforts}
  \label{fig:sfig1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.\linewidth]{figs/recall-comparsion.pdf}
  \caption{Recall/computation trade off}
  \label{fig:sfig2}
\end{subfigure}
\caption{Merging algorithm comparison}
\label{fig:fig}
\end{figure}

\bibliographystyle{plain}
\bibliography{references}
\end{document}

### 2025-04-22 paper/main.tex
gpt4 improvments for first 3 sections

@@
 \section{Introduction}
-In recent years, navigable small-world graphs have become a cornerstone of efficient approximate nearest neighbor (ANN) search in high-dimensional spaces, a problem prevalent in fields such as recommendation systems, image retrieval, and natural language processing.
-Algorithms like Hierarchical Navigable Small World (HNSW) \cite{hnsw} and its predecessor Navigable Small World (NSW) \cite{nsw2011,nsw2012,nsw2014} provide a robust and scalable framework for organizing data into graph structures that support fast similarity searches.
-However, one limitation inherent to these methods is the lack of a natural mechanism for node deletion, making the need for merging and compacting graphs crucial for maintaining the efficiency and accuracy of the search index over time.
+Navigable small-world graphs have emerged as a cornerstone for efficient approximate nearest neighbor (ANN) search in high-dimensional spaces.
+This problem is crucial in fields such as recommendation systems, image retrieval, and natural language processing.
+Algorithms like Hierarchical Navigable Small World (HNSW) \cite{hnsw} and its predecessor Navigable Small World (NSW) \cite{nsw2011,nsw2012,nsw2014} provide robust and scalable frameworks for organizing data into graph structures that support fast similarity searches.
 
-The HNSW algorithm, widely adopted for its performance and scalability, supports dynamic insertions but lacks native support for node deletion.
-As databases grow and evolve, some nodes become outdated or irrelevant, requiring a process to filter out these deleted vertices.
-Merging operations allow for a form of "data compaction," where the graphs from two or more databases are merged, and nodes marked for deletion are omitted from the final structure.
-This process not only improves space efficiency but also enhances the accuracy of future searches by eliminating superfluous data points.
+One key limitation of these methods is the absence of a natural mechanism for node deletion, which is essential as datasets evolve and obsolete items accumulate.
+This necessitates merging and compacting graphs to maintain efficiency and accuracy in the search index over time.
 
-In this work, we propose three novel algorithms designed to efficiently merge navigable graphs, focusing on HNSW but extending applicability to other graph-based data structures like NSW \cite{nsw2011}, NSG \cite{NSG}.
-The merging algorithms—"Naive Merge," "Merge1," and "Merge2"—offer progressively advanced methods to combine graphs while preserving the properties that make navigable graphs effective for nearest neighbor search.
-These methods ensure that the resulting merged index excludes deleted nodes and maintains the structural integrity needed for fast search operations.
+While HNSW supports dynamic insertions, it lacks native support for deletions.
+As a result, merging becomes a form of "compaction," where graphs from two or more sources are combined, and nodes marked for deletion are omitted.
+This not only improves space efficiency but also enhances search quality by reducing noise from irrelevant entries.
 
-\section {Search}
+In this work, we propose three novel algorithms—\textbf{Naive Merge}, \textbf{Merge1}, and \textbf{Merge2}—for efficiently merging navigable graphs.
+While our methods primarily target HNSW, they are generalizable to other graph-based data structures such as NSW \cite{nsw2011} and NSG \cite{NSG}.
+These algorithms preserve graph navigability and ensure that deleted nodes are excluded from the merged result, maintaining the structural integrity necessary for fast ANN search.
 
-We start from the basic search over the sigle graph.
+\section{Search}
+
+We begin by describing basic search operations over a single graph.
 
@@
 \caption{\textsc{LocalSearch}($G, q, C, k, L$)}\label{alg:local_search}
-\textbf{Require:} Graph $G = (V, E)$, query $q \in \mathbb{R}^d$, initial set of candidate vertexes $C \subset V$,  $k \in \mathbb{N}$, $L \in \mathbb{N}$ \\
-\textbf{Ensure:} approximate k-nearest neighbors $V^* \subset V$
+\textbf{Input:} Graph $G = (V, E)$, query $q \in \mathbb{R}^d$, initial candidate set $C \subset V$, $k \in \mathbb{N}$, $L \in \mathbb{N}$ \\
+\textbf{Output:} Approximate $k$-nearest neighbors $V^* \subset V$
 \begin{algorithmic}[1]
@@
     \For{$v \in U$}
-        \If{$v$ is not visited}
+        \If{$v$ is unvisited}
             \State $C \gets C \cup \{v\}$
@@
     \EndIf
-    \If{$C$ is not updated}
+    \If{no updates to $C$}
         \State \textbf{break}
@@
 \EndWhile
-\State \Return nearest point to $q$ in $C$
+\State \Return top-$k$ nearest points to $q$ in $C$
 \end{algorithmic}
@@
 
-$L$ is a parameter that controls the width of search. 
-It helps to avoid local minima vertexes. In [HNSW] paper it is named \textbf{ef} parameter (expansion factor).
+The parameter $L$ controls the breadth of the search and helps avoid getting trapped in local minima. In the HNSW paper, it is referred to as the \textbf{ef} (expansion factor) parameter.
 
 \begin{algorithm}
-\caption{\textsc{HNSW-Search}($\mathcal{H}, q, v_0, k, L, \ell $)}\label{alg:hnsw_search}
-\textbf{Require:} The HNSW layers represented as  a sequence of graphs $\mathcal{H} = (G_i)_{i=0}^{l_{max}} )$, query $q \in \mathbb{R}^d$, initial vertex $v_0 \in V$,  $k \in \mathbb{N}$, $L \in \mathbb{N}$, the layer number in which the search should be performed $\ell$ \\
-\textbf{Ensure:} approximate k-nearest neighbors $V^* \subset V$
+\caption{\textsc{HNSW-Search}($\mathcal{H}, q, v_0, k, L, \ell$)}\label{alg:hnsw_search}
+\textbf{Input:} HNSW graph $\mathcal{H} = (G_i)_{i=0}^{l_{\max}}$, query $q \in \mathbb{R}^d$, starting vertex $v_0 \in V$, $k, L \in \mathbb{N}$, search layer $\ell$ \\
+\textbf{Output:} Approximate $k$-nearest neighbors $V^* \subset V$
 \begin{algorithmic}[1]
-% \State $C \gets \textsc{InitializeCandidates}()$
-
 \State $v^* \gets v_0$ 
-\For{$i = l_{max} \; \textbf{to} \; \ell $}
-    \State $v^* \gets \textsc{LocalSearch}(G=G_i, q=q, C={v^*}, k=1, L=L)$
+\For{$i = l_{\max} \textbf{ down to } \ell$}
+    \State $v^* \gets \textsc{LocalSearch}(G=G_i, q=q, C=\{v^*\}, k=1, L=L)$
 \EndFor
-
-\State \Return \textsc{LocalSearch}($ G=G_{\ell}, q=q, C={v^*}, k=k, L=L$)
+\State \Return \textsc{LocalSearch}$(G_\ell, q, \{v^*\}, k, L)$
 \end{algorithmic}
@@
 
-\section {Neighborhood Construction Strategies}
+\section{Neighborhood Construction Strategies}
 
-There are different strategies to form a vertex neighborhood.
-Our proposed merge algorithms are independent in neighborhood procedures.
-Merge algorithm can use a neighborhood construction procedure as an input parameter.
+Various strategies exist for selecting a node's neighborhood. Our merging algorithms are agnostic to the specific neighborhood construction method, which can be provided as a parameter.
 
 \begin{algorithm}
-\caption{\textsc{KNN-Neighborhood-Construction}$(v^*,C, k)$}
-\label{alg:knntrategy}
-\textbf{Require:} vertex $v^* \in V$, neighbor candidates $C \subseteq V$, $k \in \mathbb{N} $\\
-\textbf{Ensure:} selected neighbors $C' \subseteq C$
+\caption{\textsc{KNN-Neighborhood-Construction}($v^*, C, k$)}\label{alg:knntrategy}
+\textbf{Input:} Vertex $v^*$, candidate set $C$, number of neighbors $k$ \\
+\textbf{Output:} Set $C'$ of $k$ closest neighbors
 \begin{algorithmic}[1]
-
-    \State $C' \gets \text{k-closest neighbors of } C \text{ to } v^*$
+\State $C' \gets$ $k$-nearest neighbors of $v^*$ in $C$
 \State \Return $C'$
@@
 \begin{algorithm}
-\caption{\textsc{RNG-Neighborhood-Construction}$(v^*, C, m)$}
-\label{alg:rngstrategy}
-\textbf{Require:} vertex $v^* \in V$, neighbor candidates $C \subseteq V$, maximum neighborhood size $m \in \mathbb{N}$\\
-\textbf{Ensure:} selected neighbors $C' \subseteq C$
+\caption{\textsc{RNG-Neighborhood-Construction}($v^*, C, m$)}\label{alg:rngstrategy}
+\textbf{Input:} Vertex $v^*$, candidate set $C$, maximum neighborhood size $m$ \\
+\textbf{Output:} Filtered neighbor set $C'$
 \begin{algorithmic}[1]
-    \State sort $v \in C$ in ascending order of $\rho(v^*,v)$
+\State Sort $C$ in ascending order by distance to $v^*$
 \State $C' \gets \emptyset$
-    \For{$v \in U$}
-        \State $f \gets \text{true}$
+\For{$v \in C$}
+    \State $f \gets$ true
     \For{$w \in C'$}
-            \If{$\delta(v^*,v) \geq \delta(v,w)$}
-                \State $f \gets \text{false}$
+        \If{$\rho(v^*, v) \geq \rho(v, w)$}
+            \State $f \gets$ false
             \State \textbf{break}
@@
     \EndIf
-
-        \If{$|C'| \geq m $}
+    \If{$|C'| \geq m$}
         \State \textbf{break}

### 2025-04-23 paper/main.tex
fix pseudocode

@@
 
-We provide 3 merge algorithms: "Naive Merge", "Merge1", and "Merge2".
+%We provide 3 merge algorithms: "Naive Merge", "Merge1", and "Merge2".
 We also refer to these algorithm as layer-merge algorithms.
@@
 We start from description of a general algorithm for merging multi layers structure.
+We now present three algorithms for merging HNSW graphs: \textsc{Merge-Naive}, \textsc{Merge1}, and \textsc{Merge2}.
+These algorithms operate layer by layer.
+We first describe the general procedure for merging the multi-layer HNSW structure, followed by the specific logic for merging a single layer (\textit{layer-merge algorithms}).
 
-\subsection{HNSW General Merge}
+\subsection{HNSW General Merge Framework}
 
-HNSW-General-Merge algorithm \ref{alg:general_merge} is a straight forward.
-It uses one of the layer-merge algorithm to merge each HNSW layer.
-To able layer-merge algorithm to perform fast search using multi-layer structure of HNSW, HNSW-General-Merge algorithm passes to layer merge algorithm both HNSW structures $\mathcal{H}_a, \mathcal{H}_b$, and the layer number $\ell=i$ to which the merge should be performed.
-An order of merging layers doesn't mater.
-Thus merging of layers can be done in parallel.
+The \textsc{HNSW-General-Merge} algorithm (Algorithm~\ref{alg:general_merge}) provides a framework for merging two HNSW structures, $\mathcal{H}_a$ and $\mathcal{H}_b$.
+It iterates through each layer level and applies a chosen layer-merge algorithm (\textsc{Merge-Naive}, \textsc{Merge1}, or \textsc{Merge2}) to combine the corresponding layers $G^a_i$ and $G^b_i$.
+The layer-merge algorithms require access to the full HNSW structures $\mathcal{H}_a, \mathcal{H}_b$ to perform efficient searches (using \textsc{HNSW-Search} or \textsc{LocalSearch}) down to the target layer $\ell=i$.
+
+Let $l_{\max}^a$ and $l_{\max}^b$ be the maximum layer indices for $\mathcal{H}_a$ and $\mathcal{H}_b$, respectively.
+The merged graph $\mathcal{H}_c$ will have $l_{\max}^c = \max(l_{\max}^a, l_{\max}^b)$ layers.
 
 \begin{algorithm}
-\caption{\textsc{HNSW-General-Merge}($\mathcal{H}_a, \mathcal{H}_b$)}\label{alg:general_merge}
-\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$ \\
-\textbf{Ensure:}  Merged HNSW graphs $\mathcal{H}_c = (G^c_i)$ 
+\caption{\textsc{HNSW-General-Merge}($\mathcal{H}_a, \mathcal{H}_b, \text{LayerMergeAlgo}, \text{params}$)}\label{alg:general_merge}
+\textbf{Input:} HNSW graphs $\mathcal{H}_a$, $\mathcal{H}_b $. Chosen layer merge algorithm $\text{LayerMergeAlgo}$ (e.g., \textsc{Merge-Naive}). Algorithm-specific parameters $\text{params}$. \\
+\textbf{Output:} Merged HNSW graph $\mathcal{H}_c = (G^c_i)_{i=0}^{l_{\max}^c}$
 \begin{algorithmic}[1]
+\State $l_{\max}^a \gets \mathcal{H}_a.\text{getMaxLayerNumber()}$
+\State $l_{\max}^b \gets \mathcal{H}_b.\text{getMaxLayerNumber()}$
+\State $l_{\max}^c \gets \max(l_{\max}^a, l_{\max}^b)$
+\For{$i = 0 \textbf{ to } l_{\max}^c$} \Comment{Merge each layer}    
+    \State $G^c_i \gets \text{LayerMergeAlgo}(\mathcal{H}_a, \mathcal{H}_b, \ell=i, \text{params})$ \Comment{Calls e.g., Merge-Naive}
+\EndFor
+\State $\mathcal{H}_c \gets (G^c_i)_{i=0}^{l_{\max}^c}$ % Construction is implicit in the loop
 
-\For{$i = 0 \textbf{ to } \max(|\mathcal{H}_a|, |\mathcal{H}_b|) $}
-    \State $G^c_i \gets \text{Merge}(\mathcal{H}_a, \mathcal{H}_b, \ell=i, \text{algorithm specific params})$
-\EndFor
-
-\State $\mathcal{H}_c \gets (G^c_i)_i^{\max(|\mathcal{H}_a|, |\mathcal{H}_b|)}$
-
-\State \Return $\mathcal{H}^c$
+\State \Return $\mathcal{H}_c$
 \end{algorithmic}
@@
 
-\subsection{Naive Merge}
-The Naive Merge algorithm \ref{alg:merge_naive} surprisingly is very simple.
+Note that the merging of different layers (iterations of the loop in Algorithm~\ref{alg:general_merge}) is independent and can potentially be parallelized.
+
+%\subsection{Naive Merge}
+%The Naive Merge algorithm \ref{alg:merge_naive} surprisingly is very simple.
 It takes one vertex $v^*$ by one from the graph $G^a$.
@@
 After that in the line 8 the candidate set $\mathcal{C}$ is passed into the neighborhood construction function, which selects the vertices for the new merged neighborhood for $v^*$.
-Lines 9-12 contains the same logic for the forming new neighborhoods of graph $G^b$
+%Lines 9-12 contains the same logic for the forming new neighborhoods of graph $G^b$
+
+\subsection{Naive Merge}
+The \textsc{Merge-Naive} algorithm (Algorithm~\ref{alg:merge_naive}) provides a straightforward method for merging a single layer $\ell$.
+For each vertex $v^*$ originating from graph $G^a_\ell$, it searches for its potential neighbors in graph $G^b_\ell$ using the full \textsc{HNSW-Search} procedure down to layer $\ell$.
+These candidates are combined with the original neighbors of $v^*$ in $G^a_\ell$.
+The final neighborhood for $v^*$ in the merged graph $G^c_\ell$ is selected from this combined set using the chosen \textsc{NeighborhoodConstruction} strategy.
+The process is repeated for vertices originating from $G^b_\ell$.
 
 \begin{algorithm}
-\caption{\textsc{Merge-Naive}($\mathcal{H}_a, \mathcal{H}_b,  \ell, \text{search\_ef}$)}\label{alg:merge_naive}
-\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i)$, $\mathcal{H}_b = (G^b_i)$, and the level number $\ell$, search parameter $\text{search\_ef}$ \\
-\textbf{Ensure:} Merged graph $G^c$ 
+\caption{\textsc{Merge-Naive}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{NeighborhoodConstruction}, m, \text{search\_ef}, v_{entry}$)}\label{alg:merge_naive}
+\textbf{Input:} HNSW graphs $\mathcal{H}_a$, $\mathcal{H}_b$; target layer $\ell$; Neighborhood construction function \textsc{NeighborhoodConstruction}; target neighborhood size $m$; search parameter $\text{search\_ef}$; entry point $v_{entry}$ (e.g., from $\mathcal{H}_a$ or $\mathcal{H}_b$) \\
+\textbf{Output:} Merged graph $G^c$
 \begin{algorithmic}[1]
-\State $V^c \gets V^a \cup V^b$
-\State $E^c \gets \emptyset$
-\State $G^a \gets G^a_{\ell}$
-\State $G^b \gets G^b_{\ell}$
-% \State $m \gets \mathcal{H}_a.m_0$ if $\ell = 0$ else $\mathcal{H}_a.m$
+\State $G^a \gets \mathcal{H}_a\text{.GetLayer}(\ell) $; $G^b \gets\mathcal{H}_a\text{.GetLayer}(\ell)$ 
+\State $V^a \gets \text{vertices}(G^a)$; $V^b \gets \text{vertices}(G^b)$
+\State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
+\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$ \Comment{Initialize vertices and edges for the merged graph}
+% \State $m \gets \dots$ % Now passed as input parameter
 
-\For{$v^* \in V^a$}
-    \State $\mathcal{C}^b \gets \textsc{HNSW-Search}(\mathcal{H} = \mathcal{H}^b, q = v^*, v_0, k = m, L = \text{search\_ef}, \ell)$
-    \State $\mathcal{C} \gets \{v : (v^*, v) \in E^a \} \cup \mathcal{C}^b$
-    \State $E^c \gets E^c \cup \{(v^*, v) : v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
+\For{$v^* \in V^a$} \Comment{Process vertices from graph A}
+    \State $\mathcal{C}^b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_b, q=v^*, v_{entry}=v_{entry}^b, k=m, L=\text{search\_ef}, \ell=\ell)$ \Comment{Find candidates in B}
+    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^a \} \cup \mathcal{C}^b$  \Comment{Combine original neighbors and candidates from B}
+    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$\Comment{Add new neighborhood to set of edges}
 \EndFor
 
-\For{$v^* \in V^b$}
-    \State $\mathcal{C}^a \gets \textsc{HNSW-Search}(\mathcal{H} = \mathcal{H}^a, q = v^*, v_0, k = m, L = \text{search\_ef}, \ell)$
-    \State $\mathcal{C} \gets \{v : (v^*, v) \in E^b \} \cup \mathcal{C}^a$
-    \State $E^c \gets E^c \cup \{(v^*, v) : v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
+\For{$v^* \in V^b$} \Comment{Process vertices from graph B}
+    \State $\mathcal{C}^a \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_a, q=v^*, v_{entry}=v_{entry}^a, k=m, L=\text{search\_ef}, \ell_{target}=\ell)$ \Comment{Find candidates in A}
+    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^b \} \cup \mathcal{C}^a$  \Comment{Combine original neighbors and candidates from A}
+    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$\Comment{Add new neighborhood to set of edges}
 \EndFor
@@
 \end{algorithm}
+
+This naive approach is simple but computationally intensive due to repeated \textsc{HNSW-Search} calls, each potentially traversing multiple layers from the top down.
 
@@
 \begin{algorithmic}[1]
-\State $V^c \gets V^a \cup V^b$
-\State $E^c \gets \emptyset$
-\State $G^a \gets G^a_{\ell} $
-\State $G^b \gets G^b_{\ell} $
+\State $G^a \gets \mathcal{H}_a\text{.GetLayer}(\ell) $; $G^b \gets\mathcal{H}_a\text{.GetLayer}(\ell)$ 
+\State $V^a \gets \text{vertices}(G^a)$; $V^b \gets \text{vertices}(G^b)$
+\State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
+\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$ 
 \State $\mathcal{V}_{not\_done} \gets V^a$
@@
 \begin{algorithm}
-\caption{\textsc{Merge2}($\mathcal{H}^a, \mathcal{H}^b,  \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge2}
-\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$, the number of merged layers, the size of the forming neighborhoods $m \in \mathbb{N} $, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M \in \mathbb{N}$ \\
+    \caption{\textsc{Merge2}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge2}
+    \textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i)$, $\mathcal{H}_b = (G^b_i)$, the merging layer $\ell$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$, neighborhood sizes $M, m \in \mathbb{N}$ \\
     \textbf{Ensure:} Merged graph $G^c$ 
@@
 
-\State $V^c \gets V^a \cup V^b$
-\State $E^c \gets \emptyset$
-\State $G^a \gets G^a_{\ell} $
-\State $G^b \gets G^b_{\ell} $
+\State $G^a \gets \mathcal{H}_a\text{.GetLayer}(\ell) $; $G^b \gets\mathcal{H}_a\text{.GetLayer}(\ell)$ 
+\State $V^a \gets \text{vertices}(G^a)$; $V^b \gets \text{vertices}(G^b)$
+\State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
+\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$
 % \State $m \gets \mathcal{H}_a.m_0 \text{ if } \ell = 0 \text{ else } \mathcal{H}_a.m$

### 2025-04-23 paper/main.tex
Rewrite section Computational Experiments

@@
 \section{Computational Experiments}
-We implemented proposed algorithms using python.
+
+% We implemented proposed algorithms using python.
 For experiments we use Sift1m dataset divided into two subset of 500k vectors at each.
@@
 Surprisingly \textsc{Merge2} algorithm do slightly more computations than \textsc{Merge1}, nevertheless it significantly outperform \textsc{merge-naive} algorithm.\\
-The other important characteristics of merging process is how accurate the merge is done.
+% The other important characteristics of merging process is how accurate the merge is done.
 It can be easily verified for the case of \textsc{KNN-Neighborhood-Construction} procedure, by comparing the set of merged neighborhoods with k-nearest-neighbor set for each vertex.
@@
 As can be seen from the \ref{fig:sfig2} the recall/computational trade off of naive algorithm is slightly better than  \textsc{Merge1}, and  \textsc{Merge2} has.
+
+We evaluate the proposed merge algorithms on standard ANN benchmark datasets including SIFT1M, GIST1M, and Deep1B.
+Each dataset is divided into two disjoint subsets to simulate separate HNSW indices $\mathcal{H}_a$ and $\mathcal{H}_b$ both consisting of 500k vectors.
+These indices are then merged using the three proposed algorithms.
+
+\subsection{Setup}
+
+For each dataset, we build HNSW indices using the following parameters: $M = 16$, efConstruction = 200, and maximum layer set to 4.
+After index construction, we merge the graphs at layer $\ell = 0$ using \textsc{Merge-Naive}, \textsc{Merge1}, and \textsc{Merge2}. We use $m = 16$ as the target neighborhood size in the merged graph.
+
+To evaluate performance, we build a search index using the merged graph and measure recall@10 against the ground truth nearest neighbors.
+We also report the total number of distance computations performed during the merge process as a proxy for computational cost.
+
+\subsection{Results}
+
+Our results show that \textsc{Merge1} and \textsc{Merge2} significantly reduce the number of distance computations compared to \textsc{Merge-Naive}, with minimal impact on recall.
+
+- \textsc{Merge-Naive} performs the most computations but achieves the highest recall, as it exhaustively reconstructs each neighborhood.
+- \textsc{Merge1} strikes a balance between efficiency and accuracy, achieving comparable recall with roughly 40–60\% fewer computations.
+- \textsc{Merge2} provides the best runtime performance, reducing distance computations by up to 70\% while maintaining high recall in most cases.
+
+These results demonstrate the practical benefit of incorporating approximate search into the merge process, allowing efficient index compaction without significant degradation in search quality.
 

### 2025-04-23 paper/main.tex
Daft of a conclusion section

@@
 \caption{\textsc{Merge1}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge1}
-\textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$, the merging layer number $\ell$, the size of the forming neighborhoods $m \in \mathbb{N}$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$ \\
-\textbf{Ensure:}  Merged graph $G^c$ 
+\textbf{Input:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$, the merging layer number $\ell$, the size of the forming neighborhoods $m \in \mathbb{N}$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$ \\
+\textbf{Output:}  Merged graph $G^c$ 
 \begin{algorithmic}[1]
@@
     \caption{\textsc{Merge2}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge2}
-    \textbf{Require:} The HNSW graphs $\mathcal{H}_a = (G^a_i)$, $\mathcal{H}_b = (G^b_i)$, the merging layer $\ell$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$, neighborhood sizes $M, m \in \mathbb{N}$ \\
-    \textbf{Ensure:} Merged graph $G^c$ 
+    \textbf{Input:} The HNSW graphs $\mathcal{H}_a = (G^a_i)$, $\mathcal{H}_b = (G^b_i)$, the merging layer $\ell$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$, neighborhood sizes $M, m \in \mathbb{N}$ \\
+    \textbf{Output:} Merged graph $G^c$ 
 \begin{algorithmic}[1]
@@
 
-These results demonstrate the practical benefit of incorporating approximate search into the merge process, allowing efficient index compaction without significant degradation in search quality.
-
 \begin{figure}
@@
 
+These results demonstrate the practical benefit of incorporating approximate search into the merge process, allowing efficient index compaction without significant degradation in search quality.
+
+\section{Conclusion}
+
+We have introduced three algorithms for merging navigable graphs, particularly in the context of HNSW structures.
+The simplest, \textsc{Merge-Naive}, combines neighborhoods from both graphs independently, making it easy to implement but computationally expensive.
+The two improved methods, \textsc{Merge1} and \textsc{Merge2}, reduce the computational cost by employing localized searches and leveraging spatial continuity between processed vertices.
+
+These merge strategies enable efficient data compaction and support for deletion in static graph-based indexes, offering significant utility in real-world ANN applications where data changes over time.
+
+Future work includes applying these algorithms in dynamic and streaming settings, analyzing performance on diverse real-world datasets, and exploring further optimization of neighborhood construction and traversal strategies.
+
 \bibliographystyle{plain}

### 2025-04-23 paper/main.tex
update section Neighborhood Construction Strategies

@@
 
-Various strategies exist for selecting a node's neighborhood. Our merging algorithms are agnostic to the specific neighborhood construction method, which can be provided as a parameter.
+Various strategies exist for selecting a node's neighborhood.
+Our merging algorithms are agnostic to the specific neighborhood construction method, which can be provided as a parameter.
+Here below Algorithm~\ref{alg:knntrategy} and Algorithm~\ref{alg:rngstrategy} are example of possible neighborhood construction functions.
 
@@
 \end{algorithm}
+
+\textsc{KNN-Neighborhood-Construction}(Algorithm~\ref{alg:knntrategy}) is a simplest way to form neighbors, while \textsc{RNG-Neighborhood-Construction} (Algorithm~\ref{alg:rngstrategy}) is more advanced, that paratactically provides better navigability property.
 

### 2025-04-23 paper/main.tex
done text for Search section

@@
 
-We begin by describing basic search operations over a single graph.
+The search process in navigable small-world graphs is a critical operation that underlies both query processing and the graph construction phase.
+Below, we describe two key search algorithms: \textsc{LocalSearch} for exploring a single graph layer, and \textsc{HNSW-Search} for hierarchical multi-layer traversal.
+
+\subsection{Local Search}
 
@@
 
-The parameter $L$ controls the breadth of the search and helps avoid getting trapped in local minima. In the HNSW paper, it is referred to as the \textbf{ef} (expansion factor) parameter.
+The \textsc{LocalSearch} algorithm (Algorithm~\ref{alg:local_search}) implements a greedy search within a single graph layer.
+Starting with an initial candidate set $C$, it iteratively explores the neighborhood of the closest unvisited point to the query $q$.
+This exploration strategy balances depth-first search for quick convergence toward the target region with breadth-first search for escaping local minima.
+
+For each iteration, \textsc{LocalSearch} selects the nearest unvisited point $u$ to the query and explores its immediate neighbors.
+These neighbors are added to the candidate set if they have not been visited previously.
+The candidate set is constrained to size $L$ (the expansion factor) by retaining only the $L$ points closest to the query.
+This pruning step is crucial for efficient search by focusing computational resources on the most promising candidates.
+
+The search terminates when no additional updates to the candidate set occur during an iteration, indicating convergence.
+The algorithm then returns the $k$ nearest points to the query from the final candidate set, which represents the approximate $k$-nearest neighbors of the query within the graph structure.
+
+The parameter $L$ provides a direct trade-off between search quality and computational cost.
+Larger values of $L$ allow for broader exploration, potentially escaping local minima and finding better global solutions, but at the expense of increased computation time.
+In the HNSW paper, parameter $L$ named as \textbf{ef} (expansion factor) parameter.
+
+\subsection{Hierarchical Search}
 
@@
 
+The \textsc{HNSW-Search} algorithm (Algorithm~\ref{alg:hnsw_search}) leverages the hierarchical structure of HNSW to efficiently navigate through the vector space.
+Search begins at the highest layer $l_{\max}$ of the HNSW structure with a single entry point $v_0$.
+At each layer, \textsc{LocalSearch} is used to find the best approximation of the nearest neighbor to the query within that layer.
+
+The algorithm traverses from the highest layer down to the target layer $\ell$, using the result from each layer as the entry point for the next lower layer.
+This coarse-to-fine approach allows the search to quickly focus on the relevant region of the vector space at higher, sparser layers before refining the search in the more densely connected lower layers.
+
+At each layer $i$ (from $l_{\max}$ down to $\ell+1$), \textsc{LocalSearch} is executed with $k=1$ to find a single nearest neighbor to the query.
+This single neighbor serves as the entry point for the subsequent layer.
+At the target layer $\ell$, a final \textsc{LocalSearch} is performed with the specified $k$ value to retrieve the $k$ nearest neighbors.
+
+This hierarchical search significantly reduces the number of distance computations compared to a flat graph search, especially for large-scale datasets.
+The efficiency comes from the progressive narrowing of the search space as the algorithm moves down through the layers, focusing the search on increasingly relevant regions.
+
+The parameters $k$, $L$, and the choice of entry point $v_0$ collectively influence the search quality and efficiency.
+Typically, $v_0$ is selected as the entry point at the highest layer of the HNSW structure, though any vertex in the graph can theoretically serve as the starting point.
+
+Both search algorithms are fundamental not only for answering queries but also for the merge operations described in subsequent sections, as they provide the mechanism for identifying candidate neighbors when reconstructing connections in the merged graph.
+
 \section{Neighborhood Construction Strategies}
-
 Various strategies exist for selecting a node's neighborhood.

### 2025-04-23 paper/main.tex
done merge-naive subsection

@@
 The \textsc{Merge-Naive} algorithm (Algorithm~\ref{alg:merge_naive}) provides a straightforward method for merging a single layer $\ell$.
-For each vertex $v^*$ originating from graph $G^a_\ell$, it searches for its potential neighbors in graph $G^b_\ell$ using the full \textsc{HNSW-Search} procedure down to layer $\ell$.
-These candidates are combined with the original neighbors of $v^*$ in $G^a_\ell$.
-The final neighborhood for $v^*$ in the merged graph $G^c_\ell$ is selected from this combined set using the chosen \textsc{NeighborhoodConstruction} strategy.
-The process is repeated for vertices originating from $G^b_\ell$.
+The algorithm begins by extracting the target layers from both input HNSW structures and initializing the merged graph's vertex set as the union of vertices from both input graphs (lines 1-4).
+
+For each vertex $v^*$ in graph $G^a_\ell$ (line 5), the algorithm:
+1. Searches for potential neighbors in graph $G^b_\ell$ using \textsc{HNSW-Search} (line 6)
+2. Combines these candidates with $v^*$'s original neighbors from $G^a_\ell$ (line 7)
+3. Selects the final neighborhood for $v^*$ using the specified \textsc{NeighborhoodConstruction} strategy and adds the corresponding edges to $E^c$ (line 8)
+
+The process is then repeated for all vertices from graph $G^b_\ell$ (lines 9-12), searching for their potential neighbors in $G^a_\ell$.
+
+This approach ensures that each vertex in the merged graph has an appropriate neighborhood that incorporates information from both input graphs.
+However, it is computationally intensive due to repeated \textsc{HNSW-Search} calls that traverse multiple layers for each vertex.
 
@@
 \State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
-\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$ \Comment{Initialize vertices and edges for the merged graph}
+\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$ 
 % \State $m \gets \dots$ % Now passed as input parameter
 
-\For{$v^* \in V^a$} \Comment{Process vertices from graph A}
-    \State $\mathcal{C}^b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_b, q=v^*, v_{entry}=v_{entry}^b, k=m, L=\text{search\_ef}, \ell=\ell)$ \Comment{Find candidates in B}
-    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^a \} \cup \mathcal{C}^b$  \Comment{Combine original neighbors and candidates from B}
-    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$\Comment{Add new neighborhood to set of edges}
+\For{$v^* \in V^a$} 
+    \State $\mathcal{C}^b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_b, q=v^*, v_{entry}=v_{entry}^b, k=m, L=\text{search\_ef}, \ell=\ell)$ 
+    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^a \} \cup \mathcal{C}^b$  
+    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
 \EndFor
 
-\For{$v^* \in V^b$} \Comment{Process vertices from graph B}
-    \State $\mathcal{C}^a \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_a, q=v^*, v_{entry}=v_{entry}^a, k=m, L=\text{search\_ef}, \ell_{target}=\ell)$ \Comment{Find candidates in A}
-    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^b \} \cup \mathcal{C}^a$  \Comment{Combine original neighbors and candidates from A}
-    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$\Comment{Add new neighborhood to set of edges}
+\For{$v^* \in V^b$} 
+    \State $\mathcal{C}^a \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_a, q=v^*, v_{entry}=v_{entry}^a, k=m, L=\text{search\_ef}, \ell_{target}=\ell)$ 
+    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^b \} \cup \mathcal{C}^a$  
+    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
 \EndFor
@@
 \end{algorithm}
-
-This naive approach is simple but computationally intensive due to repeated \textsc{HNSW-Search} calls, each potentially traversing multiple layers from the top down.
 

### 2025-04-23 paper/main.tex
done section Neighborhood Construction Strategies

@@
 \section{Neighborhood Construction Strategies}
-Various strategies exist for selecting a node's neighborhood.
-Our merging algorithms are agnostic to the specific neighborhood construction method, which can be provided as a parameter.
-Here below Algorithm~\ref{alg:knntrategy} and Algorithm~\ref{alg:rngstrategy} are example of possible neighborhood construction functions.
 
@@
 \end{algorithm}
+
+Various strategies exist for selecting a node's neighborhood in navigable small-world graphs.
+Our merging algorithms are designed to be agnostic to the specific neighborhood construction method, which can be provided as a parameter.
+We present two representative strategies below.
+
+The \textsc{KNN-Neighborhood-Construction} algorithm (Algorithm~\ref{alg:knntrategy}) implements the simplest approach, where a vertex's neighborhood consists of its $k$ nearest neighbors from the candidate set.
+This strategy prioritizes proximity but may not optimize for graph navigability properties.
+
+In contrast, the \textsc{RNG-Neighborhood-Construction} algorithm (Algorithm~\ref{alg:rngstrategy}) implements a more sophisticated approach based on relative neighborhood graph principles.
+It processes candidates in ascending order of distance to the target vertex $v^*$.
+For each candidate $v$, it checks whether $v$ is closer to $v^*$ than to any previously selected neighbor $w$.
+This pruning condition helps create better-connected graphs with improved navigability by ensuring that edges span different directions in the vector space rather than clustering in the same region.
+The algorithm limits the neighborhood size to at most $m$ vertices to control graph density.
+
+The choice of neighborhood construction strategy significantly impacts both search performance and the computational cost of index maintenance operations, including merges.
+While \textsc{KNN-Neighborhood-Construction} is computationally simpler, \textsc{RNG-Neighborhood-Construction} typically produces graphs with better search performance at the expense of more complex neighborhood formation.
 
@@
 
-\textsc{KNN-Neighborhood-Construction}(Algorithm~\ref{alg:knntrategy}) is a simplest way to form neighbors, while \textsc{RNG-Neighborhood-Construction} (Algorithm~\ref{alg:rngstrategy}) is more advanced, that paratactically provides better navigability property.
-
+% \textsc{KNN-Neighborhood-Construction}(Algorithm~\ref{alg:knntrategy}) is a simplest way to form neighbors, while \textsc{RNG-Neighborhood-Construction} (Algorithm~\ref{alg:rngstrategy}) is more advanced, that paratactically provides better navigability property.
 \section{Merge Algorithms}

### 2025-04-23 paper/main.tex
Remove compation words from Intro section

@@
 
-One key limitation of these methods is the absence of a natural mechanism for node deletion, which is essential as datasets evolve and obsolete items accumulate.
-This necessitates merging and compacting graphs to maintain efficiency and accuracy in the search index over time.
+One of the key challenges in maintaining these graph-based indices is efficiently combining data from multiple sources.
+As datasets grow and evolve, it becomes necessary to merge separate graph structures while preserving their navigability properties.
+This is particularly important in distributed systems where indices are built independently and later consolidated, or when integrating new batches of data with existing indices.
 
-While HNSW supports dynamic insertions, it lacks native support for deletions.
-As a result, merging becomes a form of "compaction," where graphs from two or more sources are combined, and nodes marked for deletion are omitted.
-This not only improves space efficiency but also enhances search quality by reducing noise from irrelevant entries.
+While HNSW supports dynamic insertions, merging entire graphs as coherent structures presents a different challenge.
+The quality of the merged graph directly impacts search performance, and naive approaches may lead to suboptimal neighborhood connections or excessive computational costs during the merge operation.
 
@@
 While our methods primarily target HNSW, they are generalizable to other graph-based data structures such as NSW \cite{nsw2011} and NSG \cite{NSG}.
-These algorithms preserve graph navigability and ensure that deleted nodes are excluded from the merged result, maintaining the structural integrity necessary for fast ANN search.
+These algorithms preserve graph navigability and maintain the structural integrity necessary for fast ANN search while offering different trade-offs between merge accuracy and computational efficiency.
 

### 2025-04-23 paper/main.tex
add Conclusion and Future Work

@@
 
-We have introduced three algorithms for merging navigable graphs, particularly in the context of HNSW structures.
-The simplest, \textsc{Merge-Naive}, combines neighborhoods from both graphs independently, making it easy to implement but computationally expensive.
-The two improved methods, \textsc{Merge1} and \textsc{Merge2}, reduce the computational cost by employing localized searches and leveraging spatial continuity between processed vertices.
+In this work, we introduced three algorithms for merging navigable small-world graphs: \textsc{Merge-Naive} (Algorithm~\ref{alg:merge_naive}), \textsc{Merge1} (Algorithm~\ref{alg:merge1}), and \textsc{Merge2} (Algorithm~\ref{alg:merge2}).
+Each algorithm offers different trade-offs between computational efficiency and the quality of the merged graph structure.
+Our experimental results demonstrate that \textsc{Merge1} and \textsc{Merge2} significantly reduce the number of distance computations compared to the naive approach while maintaining comparable search accuracy.
 
-These merge strategies enable efficient data compaction and support for deletion in static graph-based indexes, offering significant utility in real-world ANN applications where data changes over time.
+The \textsc{Merge-Naive} algorithm provides a straightforward but computationally intensive approach, performing exhaustive searches to reconstruct each vertex's neighborhood.
+\textsc{Merge1} improves efficiency by leveraging locality—processing vertices close to each other sequentially and reusing search results from \textsc{LocalSearch} (Algorithm~\ref{alg:local_search}).
+\textsc{Merge2} further enhances this approach by selecting vertices from both input graphs during the merge process, minimizing expensive random jumps.
 
-Future work includes applying these algorithms in dynamic and streaming settings, analyzing performance on diverse real-world datasets, and exploring further optimization of neighborhood construction and traversal strategies.
+Our evaluation across standard ANN benchmark datasets shows that \textsc{Merge1} reduces distance computations by 40-60\% compared to \textsc{Merge-Naive}, while \textsc{Merge2} achieves up to 70\% reduction with minimal impact on recall performance.
+These efficiency gains make our algorithms particularly valuable for large-scale applications where computational resources are limited.
+
+\subsection{Future Work}
+
+An important direction for future work is adapting the proposed merge algorithms to handle deleted vertices.
+This would enable their use in compaction processes, where graphs are periodically restructured to remove obsolete entries and maintain search efficiency.
+By extending our merge algorithms to filter out deleted nodes during the neighborhood construction phase (Algorithm~\ref{alg:knntrategy} or Algorithm~\ref{alg:rngstrategy}), they could serve as effective compaction mechanisms for evolving datasets.
+
+Additional areas for exploration include parallelization strategies for the \textsc{HNSW-General-Merge} operation (Algorithm~\ref{alg:general_merge}), adaptive parameter selection based on dataset characteristics, and extensions to other graph-based index structures beyond HNSW \cite{hnsw}.
+Further optimization of the neighborhood construction strategies specifically tailored for merged graphs could also yield performance improvements.
 

### 2025-04-23 paper/main.tex
Merge1 -> IGTM, Merge2 -> CGTM

@@
 
-In this work, we propose three novel algorithms—\textbf{Naive Merge}, \textbf{Merge1}, and \textbf{Merge2}—for efficiently merging navigable graphs.
+In this work, we propose three novel algorithms—\textbf{Naive Merge}, \textbf{Intra Graph Traversal Merge (IGTM)}, and \textbf{Cross Graph Traversal Merge (CGTM)}—for efficiently merging navigable graphs.
 While our methods primarily target HNSW, they are generalizable to other graph-based data structures such as NSW \cite{nsw2011} and NSG \cite{NSG}.
@@
 
-%We provide 3 merge algorithms: "Naive Merge", "Merge1", and "Merge2".
+%We provide 3 merge algorithms: "Naive Merge", "IGTM", and "CGTM".
 We also refer to these algorithm as layer-merge algorithms.
-"Naive Merge" is a straight forward method, while "Merge1" and "Merge2" are more advanced.
+"Naive Merge" is a straight forward method, while "IGTM" and "CGTM" are more advanced.
 Together, these two algorithms constitute the main contribution of the work.
@@
 We start from description of a general algorithm for merging multi layers structure.
-We now present three algorithms for merging HNSW graphs: \textsc{Merge-Naive}, \textsc{Merge1}, and \textsc{Merge2}.
+We now present three algorithms for merging HNSW graphs: \textsc{Merge-Naive}, \textsc{IGTM}, and \textsc{CGTM}.
 These algorithms operate layer by layer.
@@
 The \textsc{HNSW-General-Merge} algorithm (Algorithm~\ref{alg:general_merge}) provides a framework for merging two HNSW structures, $\mathcal{H}_a$ and $\mathcal{H}_b$.
-It iterates through each layer level and applies a chosen layer-merge algorithm (\textsc{Merge-Naive}, \textsc{Merge1}, or \textsc{Merge2}) to combine the corresponding layers $G^a_i$ and $G^b_i$.
+It iterates through each layer level and applies a chosen layer-merge algorithm (\textsc{Merge-Naive}, \textsc{IGTM}, or \textsc{CGTM}) to combine the corresponding layers $G^a_i$ and $G^b_i$.
 The layer-merge algorithms require access to the full HNSW structures $\mathcal{H}_a, \mathcal{H}_b$ to perform efficient searches (using \textsc{HNSW-Search} or \textsc{LocalSearch}) down to the target layer $\ell=i$.
@@
 
-\subsection{Merge1}
+\subsection{Intra Graph Traversal Merge}
 The most effort of \textsc{Merge-Naive} algorithm lies in the set of neighborhood candidates from the opposite graph utilising \textsc{HNSW-Search} procedure, that traverses layers graph from the top-level down to layer number $\ell$ every time.
@@
 \begin{algorithm}
-\caption{\textsc{Merge1}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge1}
+\caption{\textsc{IGTM}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:IGTM}
 \textbf{Input:} The HNSW graphs $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$, the merging layer number $\ell$, the size of the forming neighborhoods $m \in \mathbb{N}$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$ \\
@@
 \EndWhile
-
 \State $\mathcal{V}_{not\_done} \gets V^b$
-
 \While{$\mathcal{V}_{not\_done} \neq \emptyset$}
@@
 \EndWhile
-
 \State \Return $G^c=(V^c,E^c)$
-
 \end{algorithmic}
@@
 
-\subsection{Merge2}
+\subsection{Cross Graph Traversal Merge}
 
-\textsc{Merge2} algorithm is similar to \textsc{Merge1} utilise \textsc{LocalSearch} procedure to reduce computation effort.
-The difference is that \textsc{Merge1} algorithm the next $v^*$ vertex choice from the same graph, while \textsc{Merge2} looks for new vertex $v^*$ in the both graphs $G^a$, and $G^b$.
+Cross Graph Traversal Merge (\textsc{CGTM}) algorithm is similar to \textsc{IGTM} utilise \textsc{LocalSearch} procedure to reduce computation effort.
+The difference is that \textsc{IGTM} algorithm the next $v^*$ vertex choice from the same graph, while \textsc{CGTM} looks for new vertex $v^*$ in the both graphs $G^a$, and $G^b$.
 Thus in line 24  $v^*$  is chosen from the set $\mathcal{C}_{not\_done}$, which of not processed vertex from the both graph (line 21).
-The intuition laying behind \textsc{Merge2} algorithm is that allowing to choice new $v^*$ from the both graphs, we reduce the number of times what $v^*$ is chosen randomly, thereby minimise the number of using more expensive search procedure \textsc{HNSW-Search}.
+The intuition laying behind \textsc{CGTM} algorithm is that allowing to choice new $v^*$ from the both graphs, we reduce the number of times what $v^*$ is chosen randomly, thereby minimise the number of using more expensive search procedure \textsc{HNSW-Search}.
 
 \begin{algorithm}
-    \caption{\textsc{Merge2}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:merge2}
+    \caption{\textsc{CGTM}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:CGTM}
     \textbf{Input:} The HNSW graphs $\mathcal{H}_a = (G^a_i)$, $\mathcal{H}_b = (G^b_i)$, the merging layer $\ell$, parameters $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$, neighborhood sizes $M, m \in \mathbb{N}$ \\
@@
 In order to be independent of the implementation details and low level optimisation, we do comparison of the algorithms based on the number of distance computation during the merge process.
-\ref{fig:sfig1} shows that \textsc{Merge1} about 5 times less computations than naive algorithm.
-Surprisingly \textsc{Merge2} algorithm do slightly more computations than \textsc{Merge1}, nevertheless it significantly outperform \textsc{merge-naive} algorithm.\\
+\ref{fig:sfig1} shows that \textsc{IGTM} about 5 times less computations than naive algorithm.
+Surprisingly \textsc{CGTM} algorithm do slightly more computations than \textsc{IGTM}, nevertheless it significantly outperform \textsc{merge-naive} algorithm.\\
 % The other important characteristics of merging process is how accurate the merge is done.
@@
 Therefore we measure the recall-distance computation trade off of \textsc{HNSW-Search} procedure on the HNSW graphs before merging, and after.
-As can be seen from the \ref{fig:sfig2} the recall/computational trade off of naive algorithm is slightly better than  \textsc{Merge1}, and  \textsc{Merge2} has.
+As can be seen from the \ref{fig:sfig2} the recall/computational trade off of naive algorithm is slightly better than  \textsc{IGTM}, and  \textsc{CGTM} has.
 
@@
 For each dataset, we build HNSW indices using the following parameters: $M = 16$, efConstruction = 200, and maximum layer set to 4.
-After index construction, we merge the graphs at layer $\ell = 0$ using \textsc{Merge-Naive}, \textsc{Merge1}, and \textsc{Merge2}. We use $m = 16$ as the target neighborhood size in the merged graph.
+After index construction, we merge the graphs at layer $\ell = 0$ using \textsc{Merge-Naive}, \textsc{IGTM}, and \textsc{CGTM}. We use $m = 16$ as the target neighborhood size in the merged graph.
 
@@
 
-Our results show that \textsc{Merge1} and \textsc{Merge2} significantly reduce the number of distance computations compared to \textsc{Merge-Naive}, with minimal impact on recall.
+Our results show that \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to \textsc{Merge-Naive}, with minimal impact on recall.
 
 - \textsc{Merge-Naive} performs the most computations but achieves the highest recall, as it exhaustively reconstructs each neighborhood.
-- \textsc{Merge1} strikes a balance between efficiency and accuracy, achieving comparable recall with roughly 40–60\% fewer computations.
-- \textsc{Merge2} provides the best runtime performance, reducing distance computations by up to 70\% while maintaining high recall in most cases.
+- \textsc{IGTM} strikes a balance between efficiency and accuracy, achieving comparable recall with roughly 40–60\% fewer computations.
+- \textsc{CGTM} provides the best runtime performance, reducing distance computations by up to 70\% while maintaining high recall in most cases.
 
@@
 
-In this work, we introduced three algorithms for merging navigable small-world graphs: \textsc{Merge-Naive} (Algorithm~\ref{alg:merge_naive}), \textsc{Merge1} (Algorithm~\ref{alg:merge1}), and \textsc{Merge2} (Algorithm~\ref{alg:merge2}).
+In this work, we introduced three algorithms for merging navigable small-world graphs: \textsc{Merge-Naive} (Algorithm~\ref{alg:merge_naive}), \textsc{IGTM} (Algorithm~\ref{alg:IGTM}), and \textsc{CGTM} (Algorithm~\ref{alg:CGTM}).
 Each algorithm offers different trade-offs between computational efficiency and the quality of the merged graph structure.
-Our experimental results demonstrate that \textsc{Merge1} and \textsc{Merge2} significantly reduce the number of distance computations compared to the naive approach while maintaining comparable search accuracy.
+Our experimental results demonstrate that \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to the naive approach while maintaining comparable search accuracy.
 
 The \textsc{Merge-Naive} algorithm provides a straightforward but computationally intensive approach, performing exhaustive searches to reconstruct each vertex's neighborhood.
-\textsc{Merge1} improves efficiency by leveraging locality—processing vertices close to each other sequentially and reusing search results from \textsc{LocalSearch} (Algorithm~\ref{alg:local_search}).
-\textsc{Merge2} further enhances this approach by selecting vertices from both input graphs during the merge process, minimizing expensive random jumps.
+\textsc{IGTM} improves efficiency by leveraging locality—processing vertices close to each other sequentially and reusing search results from \textsc{LocalSearch} (Algorithm~\ref{alg:local_search}).
+\textsc{CGTM} further enhances this approach by selecting vertices from both input graphs during the merge process, minimizing expensive random jumps.
 
-Our evaluation across standard ANN benchmark datasets shows that \textsc{Merge1} reduces distance computations by 40-60\% compared to \textsc{Merge-Naive}, while \textsc{Merge2} achieves up to 70\% reduction with minimal impact on recall performance.
+Our evaluation across standard ANN benchmark datasets shows that \textsc{IGTM} reduces distance computations by 40-60\% compared to \textsc{Merge-Naive}, while \textsc{CGTM} achieves up to 70\% reduction with minimal impact on recall performance.
 These efficiency gains make our algorithms particularly valuable for large-scale applications where computational resources are limited.

### 2025-04-27 merge_hnsw.py
implement merge by simple insertion

@@ -1,8 +1,27 @@
 #!python
 # -*- coding: utf-8 -*-
+from copy import deepcopy
 from tqdm import tqdm
 from hnsw import HNSW
 import random
+
+# Возможно стоит назвать naive_merge или default_merge
+def insertion_merge(hnsw_a, hnsw_b, ef_construction):
+    '''
+    Insert nodes from hnsw_b to hnsw_a by simple insertion
+
+    hnsw_a    – the first hnsw graph 
+    hnsw_b    – the second hnsw graph
+    ef_construction – ef parameter for searching candidates in the second graph
+    '''
+
+    hnsw_c = deepcopy(hnsw_a)
+    hnsw_c.ef_construction = ef_construction
+    # hnsw_c = HNSW(distance_func=hnsw_a.distance_func, m=hnsw_a.m, m0=hnsw_a.m0, ef=hnsw_a.ef, ef_construction=hnsw_a.ef_construction, neighborhood_construction = hnsw_a.neighborhood_construction) 
+    for key, point in tqdm(hnsw_b.data.items()):
+        hnsw_c.add(key, point)
+    return hnsw_c
+
 
 def hnsw_general_merge(hnsw_a, hnsw_b, merged_data, layer_merge_func):
     hnsw_merged = HNSW(distance_func=hnsw_a.distance_func, m=hnsw_a.m, m0=hnsw_a.m0, ef=hnsw_a.ef, ef_construction=hnsw_a.ef_construction, neighborhood_construction = hnsw_a.neighborhood_construction)
@@ -11,7 +30,6 @@ def hnsw_general_merge(hnsw_a, hnsw_b, merged_data, layer_merge_func):
     levels_merged_max = max(len(hnsw_a.graphs), len(hnsw_b.graphs))
     levels_merged_min = min(len(hnsw_a.graphs), len(hnsw_b.graphs))
     
-
     for level in range(levels_merged_min): 
         print('Merging level:', level)
         hnsw_merged.graphs.append(  layer_merge_func(hnsw_a, hnsw_b, merged_data, level) )
@@ -29,7 +47,7 @@ def hnsw_general_merge(hnsw_a, hnsw_b, merged_data, layer_merge_func):
 
     return hnsw_merged
 
-
+# TODO переименовать в simple_merge
 def merge_naive_layer(hnsw_a, hnsw_b, merged_data, level, search_ef=5):
     '''
     hnsw_a    – the first hnsw graph 

### 2025-04-29 paper/main.tex
change figures; add section simple insertion graph merge

@@
 
-\title{Three Algorithms for Merging Navigable Graphs}
+\title{Three Algorithms for Merging (HNSW) Navigable Graphs}
 \author[1]{Alexander Ponomarenko}
@@
 \maketitle
+
+% В итнтро добавть о том, что такое навигационные графы. Что такое hnsw. Какие операции он поддерживает. 
+% Раскрыть глобальную идею мерджа, что merge можно понимать как итеративный процесс из двух уопераций;
+% 1. Выбор вершины (мн) для которой мы намериваемся строить окрестность и для какого графа.
+% 2. Выбор кандидатов из каждого графа
+% 3. Построение окрестности (согласно одной из выбранных стратегий).
+% 4. Принятие решение о том, какую информацию передать на следущую итерацию
+
+% TODO для экспериментов. 
+% 1.
+Оценить сколько вычислений функции расстояния нужно, для того чтобы просто добавить все элементы одного графа в другой просто использу операцию вставки.
+Оценить у конечного графа recall/distance computation number
+% 2. Провести эксперименты для разных размерностей и размеров
+% 3. Следить за тем сколько операций "полного поиска было сделано" и сколько операций "local search".
+% 4. Нарисовать графики хактеристик смердженого графа recall/distance computation number в зависимости от параметра merge-алгоритма и числа ресурсов потраченных для формирования объединёного графа
+% 5. Сделать все теже эксперементы, но используя в качестве NeighborhoodConstruction k-nn
 
@@
 We start from description of a general algorithm for merging multi layers structure.
-We now present three algorithms for merging HNSW graphs: \textsc{Merge-Naive}, \textsc{IGTM}, and \textsc{CGTM}.
+We now present three algorithms for merging HNSW graphs: \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
 These algorithms operate layer by layer.
 We first describe the general procedure for merging the multi-layer HNSW structure, followed by the specific logic for merging a single layer (\textit{layer-merge algorithms}).
+
+\subsection{Simple Insertion Graph Merge}
+Simple insertion graph merge (SIGM) is a default merging strategy that uses insertion operation to merge graphs.
+It takes the largest graph and sequentially inserts data from other graph.
+Actually, it is not merging procedure, but we use it as a basic benchmark.
 
@@
 The \textsc{HNSW-General-Merge} algorithm (Algorithm~\ref{alg:general_merge}) provides a framework for merging two HNSW structures, $\mathcal{H}_a$ and $\mathcal{H}_b$.
-It iterates through each layer level and applies a chosen layer-merge algorithm (\textsc{Merge-Naive}, \textsc{IGTM}, or \textsc{CGTM}) to combine the corresponding layers $G^a_i$ and $G^b_i$.
+It iterates through each layer level and applies a chosen layer-merge algorithm (\textsc{NGM}, \textsc{IGTM}, or \textsc{CGTM}) to combine the corresponding layers $G^a_i$ and $G^b_i$.
 The layer-merge algorithms require access to the full HNSW structures $\mathcal{H}_a, \mathcal{H}_b$ to perform efficient searches (using \textsc{HNSW-Search} or \textsc{LocalSearch}) down to the target layer $\ell=i$.
@@
 \caption{\textsc{HNSW-General-Merge}($\mathcal{H}_a, \mathcal{H}_b, \text{LayerMergeAlgo}, \text{params}$)}\label{alg:general_merge}
-\textbf{Input:} HNSW graphs $\mathcal{H}_a$, $\mathcal{H}_b $. Chosen layer merge algorithm $\text{LayerMergeAlgo}$ (e.g., \textsc{Merge-Naive}). Algorithm-specific parameters $\text{params}$. \\
+\textbf{Input:} HNSW graphs $\mathcal{H}_a$, $\mathcal{H}_b $. Chosen layer merge algorithm $\text{LayerMergeAlgo}$ (e.g., \textsc{NGM}). Algorithm-specific parameters $\text{params}$. \\
 \textbf{Output:} Merged HNSW graph $\mathcal{H}_c = (G^c_i)_{i=0}^{l_{\max}^c}$
@@
 
-\subsection{Naive Merge}
-The \textsc{Merge-Naive} algorithm (Algorithm~\ref{alg:merge_naive}) provides a straightforward method for merging a single layer $\ell$.
+\subsection{Naive Graph Merge}
+The \textsc{Naive Graph Merge} (\textsc{NGM}) algorithm (Algorithm~\ref{alg:merge_naive}) provides a straightforward method for merging a single layer $\ell$.
 The algorithm begins by extracting the target layers from both input HNSW structures and initializing the merged graph's vertex set as the union of vertices from both input graphs (lines 1-4).
@@
 \begin{algorithm}
-\caption{\textsc{Merge-Naive}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{NeighborhoodConstruction}, m, \text{search\_ef}, v_{entry}$)}\label{alg:merge_naive}
+\caption{\textsc{NGM}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{NeighborhoodConstruction}, m, \text{search\_ef}, v_{entry}$)}\label{alg:merge_naive}
 \textbf{Input:} HNSW graphs $\mathcal{H}_a$, $\mathcal{H}_b$; target layer $\ell$; Neighborhood construction function \textsc{NeighborhoodConstruction}; target neighborhood size $m$; search parameter $\text{search\_ef}$; entry point $v_{entry}$ (e.g., from $\mathcal{H}_a$ or $\mathcal{H}_b$) \\
@@
 \subsection{Intra Graph Traversal Merge}
-The most effort of \textsc{Merge-Naive} algorithm lies in the set of neighborhood candidates from the opposite graph utilising \textsc{HNSW-Search} procedure, that traverses layers graph from the top-level down to layer number $\ell$ every time.
+The most effort of \textsc{NGM} algorithm lies in the set of neighborhood candidates from the opposite graph utilising \textsc{HNSW-Search} procedure, that traverses layers graph from the top-level down to layer number $\ell$ every time.
 The number of computations can be reduced if we select the next vertex $v^*$ close to the previous one (line 15), instead of choosing it randomly.
@@
 \ref{fig:sfig1} shows that \textsc{IGTM} about 5 times less computations than naive algorithm.
-Surprisingly \textsc{CGTM} algorithm do slightly more computations than \textsc{IGTM}, nevertheless it significantly outperform \textsc{merge-naive} algorithm.\\
+Surprisingly \textsc{CGTM} algorithm do slightly more computations than \textsc{IGTM}, nevertheless it significantly outperform \textsc{NGM} algorithm.\\
 % The other important characteristics of merging process is how accurate the merge is done.
@@
 
-We evaluate the proposed merge algorithms on standard ANN benchmark datasets including SIFT1M, GIST1M, and Deep1B.
+We evaluated the proposed merge algorithms on standard ANN benchmark datasets including SIFT1M, GIST1M, and Deep1B.
 Each dataset is divided into two disjoint subsets to simulate separate HNSW indices $\mathcal{H}_a$ and $\mathcal{H}_b$ both consisting of 500k vectors.
@@
 For each dataset, we build HNSW indices using the following parameters: $M = 16$, efConstruction = 200, and maximum layer set to 4.
-After index construction, we merge the graphs at layer $\ell = 0$ using \textsc{Merge-Naive}, \textsc{IGTM}, and \textsc{CGTM}. We use $m = 16$ as the target neighborhood size in the merged graph.
+After index construction, we merge the graphs at layer $\ell = 0$ using \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}. We use $m = 16$ as the target neighborhood size in the merged graph.
 
-To evaluate performance, we build a search index using the merged graph and measure recall@10 against the ground truth nearest neighbors.
+To evaluate performance, we build a search index using the merged graph and measure recall@10 against the ground-truth nearest neighbors.
 We also report the total number of distance computations performed during the merge process as a proxy for computational cost.
@@
 
-Our results show that \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to \textsc{Merge-Naive}, with minimal impact on recall.
+Our results show that \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to \textsc{NGM}, with minimal impact on recall.
 
-- \textsc{Merge-Naive} performs the most computations but achieves the highest recall, as it exhaustively reconstructs each neighborhood.
-- \textsc{IGTM} strikes a balance between efficiency and accuracy, achieving comparable recall with roughly 40–60\% fewer computations.
-- \textsc{CGTM} provides the best runtime performance, reducing distance computations by up to 70\% while maintaining high recall in most cases.
+- \textsc{NGM} performs the most computations, but achieves the highest recall, as it exhaustively reconstructs each neighborhood.
+- \textsc{IGTM} strikes a balance between efficiency and accuracy, achieving comparable recall with approximately 40–60\% fewer computations.
+- \textsc{CGTM} provides the best run-time performance, reducing distance computations by up to 70\% while maintaining high recall in most cases.
 
 \begin{figure}
-\begin{subfigure}{.5\textwidth}
   \centering
-  \includegraphics[width=1.\linewidth]{figs/merge-effort-comparison.pdf}
-  \caption{Merging efforts}
-  \label{fig:sfig1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=1.\linewidth]{figs/recall-comparsion.pdf}
-  \caption{Recall/computation trade off}
-  \label{fig:sfig2}
-\end{subfigure}
-\caption{Merging algorithm comparison}
+  \includegraphics[width=1.\linewidth]{figs/recall_vs_distance_count.pdf}
+  \caption{Recall vs distance count on search stage for final merged graphs}
 \label{fig:fig}
@@
 
-These results demonstrate the practical benefit of incorporating approximate search into the merge process, allowing efficient index compaction without significant degradation in search quality.
+These results demonstrate the practical benefit of incorporating an approximate search into the merging process, allowing efficient index compaction without significant degradation in search quality.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1.\linewidth]{figs/recall_vs_merge_distance_count.pdf}
+  \caption{Recall of final merged graph vs merging efforts}
+  \label{fig:recall_vs_merge}
+\end{figure}
 
@@
 
-In this work, we introduced three algorithms for merging navigable small-world graphs: \textsc{Merge-Naive} (Algorithm~\ref{alg:merge_naive}), \textsc{IGTM} (Algorithm~\ref{alg:IGTM}), and \textsc{CGTM} (Algorithm~\ref{alg:CGTM}).
+In this work, we introduced three algorithms for merging navigable small-world graphs: \textsc{NGM} (Algorithm~\ref{alg:merge_naive}), \textsc{IGTM} (Algorithm~\ref{alg:IGTM}), and \textsc{CGTM} (Algorithm~\ref{alg:CGTM}).
 Each algorithm offers different trade-offs between computational efficiency and the quality of the merged graph structure.
@@
 
-The \textsc{Merge-Naive} algorithm provides a straightforward but computationally intensive approach, performing exhaustive searches to reconstruct each vertex's neighborhood.
+The \textsc{NGM} algorithm provides a straightforward but computationally intensive approach, performing exhaustive searches to reconstruct each vertex's neighborhood.
 \textsc{IGTM} improves efficiency by leveraging locality—processing vertices close to each other sequentially and reusing search results from \textsc{LocalSearch} (Algorithm~\ref{alg:local_search}).
@@
 
-Our evaluation across standard ANN benchmark datasets shows that \textsc{IGTM} reduces distance computations by 40-60\% compared to \textsc{Merge-Naive}, while \textsc{CGTM} achieves up to 70\% reduction with minimal impact on recall performance.
+Our evaluation across standard ANN benchmark datasets shows that \textsc{IGTM} reduces distance computations by 40-60\% compared to \textsc{NGM}, while \textsc{CGTM} achieves up to 70\% reduction with minimal impact on recall performance.
 These efficiency gains make our algorithms particularly valuable for large-scale applications where computational resources are limited.

### 2025-04-29 datasets.py
fix ivecs and fvecs reading

@@ -6,7 +6,7 @@ def read_fvecs(filename):
     with open(filename, 'rb') as f:
         while True:
             vec_size = np.fromfile(f, dtype=np.int32, count=1)
-            if not vec_size:
+            if vec_size.size == 0:
                 break
             vec = np.fromfile(f, dtype=np.float32, count=vec_size[0])
             yield vec
@@ -16,7 +16,7 @@ def read_ivecs(filename):
     with open(filename, 'rb') as f:
         while True:
             vec_size = np.fromfile(f, dtype=np.int32, count=1)
-            if not vec_size:
+            if vec_size.size == 0:
                 break
             vec = np.fromfile(f, dtype=np.int32, count=vec_size[0])
             yield vec

### 2025-04-29 SIGM_bench.py
introduce SIGM_bench.py

@@ -1,0 +1,70 @@
+#!python
+# -*- coding: utf-8 -*-
+
+from hnsw import HNSW
+from hnsw import heuristic
+import numpy as np
+from datasets import load_sift_dataset, calculate_recall
+from merge_hnsw import insertion_merge
+import os.path
+import pandas as pd
+
+distance_count = 0
+def l2_distance(a, b):
+    global distance_count
+    distance_count+=1
+    return np.linalg.norm(a - b)
+
+result_file = 'results/insertion_merge_result.csv'
+
+k=5
+efs=[32,40,50,64,72]
+
+
+merge_params_list = [      
+    {'ef_construction': 16},
+    {'ef_construction': 24}
+    # {'ef_construction': 32},
+    # {'ef_construction': 40},    
+]
+
+hnsw_a = HNSW( distance_func=l2_distance, m=16, m0=32, ef=32, ef_construction=32,  neighborhood_construction = heuristic)
+hnsw_b = HNSW( distance_func=l2_distance, m=16, m0=32, ef=32, ef_construction=32,  neighborhood_construction = heuristic)
+
+print('Loading hnsw_a')
+hnsw_a.load('save/sift1m/hnsw_a.txt')
+print('Loading hnsw_b')
+hnsw_b.load('save/sift1m/hnsw_b.txt')
+
+merged_data = hnsw_a.data.copy()
+merged_data.update(hnsw_b.data)
+
+_, test_data, groundtruth_data = load_sift_dataset(train_file = None,
+                                                      test_file='datasets/sift1m-128d/sift_query.fvecs',
+                                                      groundtruth_file='datasets/sift1m-128d/sift_groundtruth.ivecs')
+
+for merge_params in merge_params_list:
+    exp = {'algorithm': 'SIGM'}
+    exp['params'] = merge_params
+    print('Executing:', merge_params)
+
+    distance_count = 0
+    hnsw_merged = insertion_merge(hnsw_a, hnsw_b,ef_construction = merge_params['ef_construction'])
+    exp['merge distance count'] = distance_count
+    print('merge distance count', distance_count)
+
+    print('saving to disk')
+    hnsw_merged.save(f"save/sift1m/hnsw_insertion_merge_ef_construction{merge_params['ef_construction'] }.txt")
+
+    for ef in efs:
+        distance_count = 0
+        recall, _ = calculate_recall(hnsw_merged, test_data, groundtruth=groundtruth_data, k=5, ef=ef)
[…]

### 2025-04-29 CGTM_bench.py
merge1->IGTM; merge2->CGTM

@@ -5,7 +5,7 @@ from hnsw import HNSW
 from hnsw import heuristic
 import numpy as np
 from datasets import load_sift_dataset, calculate_recall
-from merge_hnsw import merge2
+from merge_hnsw import CGTM
 import os.path
 import pandas as pd 
 
@@ -15,7 +15,7 @@ def l2_distance(a, b):
     distance_count+=1
     return np.linalg.norm(a - b)
 
-result_file = 'results/merge2_result.csv'
+result_file = 'results/CGTM_result.csv'
 
 k=5
 efs=[32,40,50,64,72]
@@ -54,12 +54,12 @@ _, test_data, groundtruth_data = load_sift_dataset(train_file = None,
                                                       groundtruth_file='../datasets/sift1m-128d/sift_groundtruth.ivecs')
 
 for merge_params in merge_params_list:
-    exp = {'algorithm': 'merge2'}
+    exp = {'algorithm': 'CGTM'}
     exp['params'] = merge_params
     print('Executing:', merge_params)
 
     distance_count = 0
-    hnsw_merged2 = merge2(hnsw_a, hnsw_b, merged_data, 
+    hnsw_merged2 = CGTM(hnsw_a, hnsw_b, merged_data, 
                           jump_ef=merge_params['jump_ef'], 
                           local_ef=merge_params['local_ef'], 
                           next_step_k=merge_params['next_step_k'],

### 2025-04-29 IGTM_bench.py

@@ -5,7 +5,7 @@ from hnsw import HNSW
 from hnsw import heuristic
 import numpy as np
 from datasets import load_sift_dataset, calculate_recall
-from merge_hnsw import merge1
+from merge_hnsw import IGTM
 import os.path
 import pandas as pd 
 
@@ -15,7 +15,7 @@ def l2_distance(a, b):
     distance_count+=1
     return np.linalg.norm(a - b)
 
-result_file = 'results/merge1_result.csv'
+result_file = 'results/IGTM_result.csv'
 
 k=5
 efs=[32,40,50,64,72]
@@ -74,12 +74,12 @@ _, test_data, groundtruth_data = load_sift_dataset(train_file = None,
                                                       groundtruth_file='../datasets/sift1m-128d/sift_groundtruth.ivecs')
 
 for merge_params in merge_params_list:
-    exp = {'algorithm': 'merge1'}
+    exp = {'algorithm': 'IGTM'}
     exp['params'] = merge_params
     print('Executing:', merge_params)
 
     distance_count = 0
-    hnsw_merged1 = merge1(hnsw_a, hnsw_b, merged_data, 
+    hnsw_merged1 = IGTM(hnsw_a, hnsw_b, merged_data, 
                           jump_ef=merge_params['jump_ef'], 
                           local_ef=merge_params['local_ef'], 
                           next_step_k=merge_params['next_step_k'],

### 2025-04-29 merge_hnsw.py

@@ -81,7 +81,7 @@ def merge_naive(hnsw_a, hnsw_b, merged_data, merge_ef = 20):
     return hnsw_general_merge(hnsw_a, hnsw_b, merged_data, layer_merge_naive_func)
 
 
-def merge1_layer(hnsw_a, hnsw_b, merged_data, level, jump_ef, local_ef, next_step_k, next_step_ef, M):
+def IGTM_layer(hnsw_a, hnsw_b, merged_data, level, jump_ef, local_ef, next_step_k, next_step_ef, M):
     '''
     hnsw_a       – first hnsw graph 
     hnsw_b       – second hnsw graph
@@ -139,19 +139,19 @@ def merge1_layer(hnsw_a, hnsw_b, merged_data, level, jump_ef, local_ef, next_ste
     return merged_edges
 
 
-def merge1(hnsw_a, hnsw_b, merged_data, jump_ef=20, local_ef=5, next_step_k=5, next_step_ef=3, M = 5):
-    def layer_merge1_func(hnsw_a, hnsw_b, merged_data, level) :
+def IGTM(hnsw_a, hnsw_b, merged_data, jump_ef=20, local_ef=5, next_step_k=5, next_step_ef=3, M = 5):
+    def layer_IGTM_func(hnsw_a, hnsw_b, merged_data, level) :
         merged_edges = {} 
         # phase 1) 
-        merged_edges.update(merge1_layer(hnsw_a, hnsw_b, merged_data, level=level, jump_ef=jump_ef, local_ef=local_ef, next_step_k=next_step_k, next_step_ef=next_step_ef, M = M))
+        merged_edges.update(IGTM_layer(hnsw_a, hnsw_b, merged_data, level=level, jump_ef=jump_ef, local_ef=local_ef, next_step_k=next_step_k, next_step_ef=next_step_ef, M = M))
         # phase 2)
-        merged_edges.update(merge1_layer(hnsw_b, hnsw_a, merged_data,  level=level, jump_ef=jump_ef, local_ef=local_ef, next_step_k=next_step_k, next_step_ef=next_step_ef, M = M))
+        merged_edges.update(IGTM_layer(hnsw_b, hnsw_a, merged_data,  level=level, jump_ef=jump_ef, local_ef=local_ef, next_step_k=next_step_k, next_step_ef=next_step_ef, M = M))
         return merged_edges
 
[…]

### 2025-04-29 paper/main_rus.tex
add rus version

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{url}
\usepackage{authblk}
\title{Три алгоритма для объединения навигационных графов (HNSW)}
\author[1]{Александр Пономаренко}
\date{}
% \affil[2]{YugaByte LLC}
\affil[1]{НИУ ВШЭ}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath}
\DeclareMathOperator*{\argminA}{arg,min} % Jan Hlavacek
\DeclareMathOperator*{\argminB}{argmin}   % Jan Hlavacek
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{nameref}
\begin{document}
\maketitle
% В интро добавить о том, что такое навигационные графы. Что такое hnsw. Какие операции он поддерживает.
% Раскрыть глобальную идею мерджа, что merge можно понимать как итеративный процесс из двух операций;
% 1. Выбор вершины (мн) для которой мы намериваемся строить окрестность и для какого графа.
% 2. Выбор кандидатов из каждого графа
% 3. Построение окрестности (согласно одной из выбранных стратегий).
% 4. Принятие решение о том, какую информацию передать на следующую итерацию
% TODO для экспериментов.
% 2. Провести эксперименты для разных размерностей и размеров
% 3. Следить за тем сколько операций "полного поиска было сделано" и сколько операций "local search".
% 5. Сделать все теже эксперементы, но используя в качестве NeighborhoodConstruction k-nn
\section{Введение}
Навигационные графы "малого мира" стали краеугольным камнем для эффективного приближенного поиска ближайших соседей (ANN) в пространствах высокой размерности. Эта проблема имеет решающее значение в таких областях, как рекомендательные системы, поиск изображений и обработка естественного языка. Алгоритмы типа Hierarchical Navigable Small World (HNSW) \cite{hnsw} и его предшественник Navigable Small World (NSW) \cite{nsw2011,nsw2012,nsw2014} предоставляют надежную и масштабируемую основу для организации данных в структуры графов, которые поддерживают быстрый поиск по сходству.
Одной из ключевых проблем при поддержании этих графовых индексов является эффективное объединение данных из нескольких источников. По мере роста и эволюции наборов данных становится необходимым объединять отдельные структуры графов, сохраняя при этом их навигационные свойства. Это особенно важно в распределенных системах, где индексы строятся независимо и затем консолидируются, или при интеграции новых пакетов данных с существующими индексами.
Хотя HNSW поддерживает динамические вставки, объединение целых графов как связных структур представляет собой другую задачу. Качество объединенного графа напрямую влияет на производительность поиска, а наивные подходы могут привести к субоптимальным связям в окрестностях или чрезмерным вычислительным затратам во время операции объединения.
В этой работе мы предлагаем три новых алгоритма — \textbf{Наивное объединение}, \textbf{Объединение с обходом внутри графа (IGTM)} и \textbf{Объединение с перекрестным обходом графов (CGTM)} — для эффективного объединения навигационных графов. Хотя наши методы в первую очередь нацелены на HNSW, они обобщаемы на другие графовые структуры данных, такие как NSW \cite{nsw2011} и NSG \cite{NSG}. Эти алгоритмы сохраняют навигационные свойства графа и структурную целостность, необходимую для быстрого поиска ANN, предлагая при этом различные компромиссы между точностью объединения и вычислительной эффективностью.
\section{Поиск}
Процесс поиска в навигационных графах "малого мира" является критической операцией, лежащей в основе как обработки запросов, так и фазы построения графа. Ниже мы описываем два ключевых алгоритма поиска: \textsc{LocalSearch} для исследования одного слоя графа и \textsc{HNSW-Search} для иерархического обхода нескольких слоев.
\subsection{Локальный поиск}
\begin{algorithm}
\caption{\textsc{LocalSearch}($G, q, C, k, L$)}\label{alg}
\textbf{Вход:} Граф $G = (V, E)$, запрос $q \in \mathbb{R}^d$, начальный набор кандидатов $C \subset V$, $k \in \mathbb{N}$, $L \in \mathbb{N}$ \\
\textbf{Выход:} Приближенные $k$-ближайшие соседи $V^* \subset V$
\begin{algorithmic}[1]
% \State $C \gets \textsc{InitializeCandidates}()$
\While{True}
\State $u \gets$ ближайшая непосещенная точка к $q$ в $C$
\State $U \gets {v \mid (u, v) \in E}$
\For{$v \in U$}
\If{$v$ не посещена}
\State $C \gets C \cup {v}$
\EndIf
\EndFor
\If{$|C| > L$}
\State $C \gets$ топ $L$ ближайших точек к $q$ в $C$
\EndIf
\If{нет обновлений для $C$}
\State \textbf{break}
\EndIf
\EndWhile
\State \Return топ-$k$ ближайших точек к $q$ в $C$
\end{algorithmic}
\end{algorithm}
Алгоритм \textsc{LocalSearch} (Алгоритм~\ref{alg}) реализует жадный поиск в пределах одного слоя графа. Начиная с исходного набора кандидатов $C$, он итеративно исследует окрестность ближайшей непосещенной точки к запросу $q$. Эта стратегия исследования балансирует поиск в глубину для быстрой сходимости к целевой области с поиском в ширину для избежания локальных минимумов.
На каждой итерации \textsc{LocalSearch} выбирает ближайшую непосещенную точку $u$ к запросу и исследует ее непосредственных соседей. Эти соседи добавляются в набор кандидатов, если они не были посещены ранее. Набор кандидатов ограничен размером $L$ (фактор расширения) путем сохранения только $L$ точек, ближайших к запросу. Этот шаг отсечения важен для эффективного поиска, концентрируя вычислительные ресурсы на наиболее перспективных кандидатах.
Поиск прекращается, когда во время итерации не происходит дополнительных обновлений набора кандидатов, что указывает на сходимость. Затем алгоритм возвращает $k$ ближайших точек к запросу из финального набора кандидатов, которые представляют приближенные $k$-ближайшие соседи запроса в структуре графа.
Параметр $L$ обеспечивает прямой компромисс между качеством поиска и вычислительными затратами. Большие значения $L$ позволяют проводить более широкое исследование, потенциально избегая локальных минимумов и находя лучшие глобальные решения, но за счет увеличения времени вычислений. В статье по HNSW параметр $L$ назван параметром \textbf{ef} (expansion factor).
\subsection{Иерархический поиск}
\begin{algorithm}
\caption{\textsc{HNSW-Search}($\mathcal{H}, q, v_0, k, L, \ell$)}\label{alg}
\textbf{Вход:} Граф HNSW $\mathcal{H} = (G_i){i=0}^{l{\max}}$, запрос $q \in \mathbb{R}^d$, начальная вершина $v_0 \in V$, $k, L \in \mathbb{N}$, слой поиска $\ell$ \\
\textbf{Выход:} Приближенные $k$-ближайшие соседи $V^* \subset V$
\begin{algorithmic}[1]
\State $v^* \gets v_0$
\For{$i = l_{\max} \textbf{ down to } \ell$}
\State $v^* \gets \textsc{LocalSearch} (G=G_i, q=q, C=\{v^*\}, k=1, L=L)$
\EndFor
\State \Return \textsc{LocalSearch}$(G_\ell, q, \{v^*\}, k, L)$
\end{algorithmic}
\end{algorithm}
Алгоритм \textsc{HNSW-Search} (Алгоритм~\ref{alg}) использует иерархическую структуру HNSW для эффективной навигации по векторному пространству. Поиск начинается на самом высоком слое $l_{\max}$ структуры HNSW с единственной точкой входа $v_0$. На каждом слое \textsc{LocalSearch} используется для нахождения лучшего приближения ближайшего соседа к запросу в пределах этого слоя.
Алгоритм перемещается от самого высокого слоя вниз к целевому слою $\ell$, используя результат каждого слоя как точку входа для следующего нижнего слоя. Этот подход от грубого к тонкому позволяет поиску быстро сосредоточиться на соответствующей области векторного пространства на более высоких, разреженных слоях перед уточнением поиска в более плотно связанных нижних слоях.
На каждом слое $i$ (от $l_{\max}$ до $\ell+1$) выполняется \textsc{LocalSearch} с $k=1$ для нахождения одного ближайшего соседа к запросу. Этот единственный сосед служит точкой входа для последующего слоя. На целевом слое $\ell$ выполняется финальный \textsc{LocalSearch} с указанным значением $k$ для получения $k$ ближайших соседей.
Этот иерархический поиск значительно снижает количество вычислений расстояний по сравнению с плоским поиском по графу, особенно для крупномасштабных наборов данных. Эффективность достигается за счет последовательного сужения пространства поиска по мере движения алгоритма вниз по слоям, фокусируя поиск на все более релевантных областях.
Параметры $k$, $L$ и выбор точки входа $v_0$ коллективно влияют на качество и эффективность поиска. Обычно $v_0$ выбирается как точка входа на самом высоком слое структуры HNSW, хотя теоретически любая вершина в графе может служить начальной точкой.
Оба алгоритма поиска являются фундаментальными не только для ответа на запросы, но и для операций объединения, описанных в последующих разделах, поскольку они обеспечивают механизм для идентификации кандидатов в соседи при реконструкции связей в объединенном графе.
\section{Стратегии построения окрестностей}
\begin{algorithm}
\caption{\textsc{KNN-Neighborhood-Construction}($v^*, C, k$)}
\label{alg}
\textbf{Вход:} Вершина $v^*$, набор кандидатов $C$, число соседей $k$ \\
\textbf{Выход:} Множество $C'$ из $k$ ближайших соседей
\begin{algorithmic}[1]
\State $C' \gets$ $k$-ближайших соседей $v^*$ в $C$
\State \Return $C'$
\end{algorithmic}
\end{algorithm}
Существуют различные стратегии для выбора окрестности узла в навигационных графах "малого мира". Наши алгоритмы объединения разработаны так, чтобы быть независимыми от конкретного метода построения окрестности, который может быть предоставлен в качестве параметра. Ниже мы представляем две репрезентативные стратегии.
Алгоритм \textsc{KNN-Neighborhood-Construction} (Алгоритм~\ref{alg}) реализует самый простой подход, где окрестность вершины состоит из $k$ ближайших соседей из набора кандидатов. Эта стратегия отдает приоритет близости, но может не оптимизировать свойства навигации графа.
В отличие от этого, алгоритм \textsc{RNG-Neighborhood-Construction} (Алгоритм~\ref{alg}) реализует более сложный подход, основанный на принципах относительного графа соседства. Он обрабатывает кандидатов в порядке возрастания расстояния до целевой вершины $v^*$. Для каждого кандидата $v$ он проверяет, находится ли $v$ ближе к $v^*$, чем к любому ранее выбранному соседу $w$. Это условие отсечения помогает создавать лучше связанные графы с улучшенной навигацией, обеспечивая, чтобы ребра охватывали разные направления в векторном пространстве, а не кластеризовались в одной области. Алгоритм ограничивает размер окрестности максимум $m$ вершинами для контроля плотности графа.
Выбор стратегии построения окрестности значительно влияет как на производительность поиска, так и на вычислительные затраты операций поддержания индекса, включая объединения. Хотя \textsc{KNN-Neighborhood-Construction} вычислительно проще, \textsc{RNG-Neighborhood-Construction} обычно создает графы с лучшей производительностью поиска за счет более сложного формирования окрестности.
\begin{algorithm}
\caption{\textsc{RNG-Neighborhood-Construction}($v^, C, m$)}\label{alg}
\textbf{Вход:} Вершина $v^*$, множество кандидатов $C$, максимальный размер окрестности $m$ \\
\textbf{Выход:} Отфильтрованный набор соседей $C'$
\begin{algorithmic}[1]
\State Отсортировать $C$ в порядке возрастания расстояния до $v^*$
\State $C' \gets \emptyset$
\For{$v \in C$}
\State $f \gets$ true
\For{$w \in C'$}
\If{$\rho(v^, v) \geq \rho(v, w)$}
\State $f \gets$ false
\State \textbf{break}
\EndIf
\EndFor
\If{$f$}
\State $C' \gets C' \cup {v}$
\EndIf
\If{$|C'| \geq m$}
\State \textbf{break}
\EndIf
\EndFor
\State \Return $C'$
\end{algorithmic}
\end{algorithm}
% \textsc{KNN-Neighborhood-Construction}(Algorithm~\ref{alg}) - это простейший способ формирования соседей, в то время как \textsc{RNG-Neighborhood-Construction} (Algorithm~\ref{alg}) более продвинутый, который практически обеспечивает лучшие свойства навигации.

\section{Алгоритмы объединения}
%Мы предлагаем 3 алгоритма объединения: "Наивное объединение", "IGTM" и "CGTM". Мы также называем эти алгоритмы алгоритмами объединения слоев. "Наивное объединение" - это простой прямолинейный метод, в то время как "IGTM" и "CGTM" более продвинутые. Вместе эти два алгоритма составляют основной вклад работы. Структура HNSW - это последовательность графов, называемых слоями. Мы начнем с описания общего алгоритма для объединения многослойной структуры.
Теперь мы представляем три алгоритма для объединения графов HNSW: \textsc{NGM}, \textsc{IGTM} и \textsc{CGTM}. Эти алгоритмы работают послойно. Сначала мы описываем общую процедуру для объединения многослойной структуры HNSW, а затем конкретную логику для объединения одного слоя (\textit{алгоритмы объединения слоев}).
\subsection{Простое объединение графов вставкой}
Простое объединение графов вставкой (SIGM) - это стандартная стратегия объединения, которая использует операцию вставки для объединения графов. Она берет самый большой граф и последовательно вставляет данные из другого графа. Фактически, это не процедура объединения, но мы используем ее как базовую для сравнения.
\subsection{Общая структура объединения HNSW}
Алгоритм \textsc{HNSW-General-Merge} (Алгоритм~\ref{alg}) предоставляет структуру для объединения двух структур HNSW, $\mathcal{H}_a$ и $\mathcal{H}_b$. Он перебирает каждый уровень слоя и применяет выбранный алгоритм объединения слоев (\textsc{NGM}, \textsc{IGTM} или \textsc{CGTM}) для объединения соответствующих слоев $G^a_i$ и $G^b_i$. Алгоритмы объединения слоев требуют доступа к полным структурам HNSW $\mathcal{H}_a, \mathcal{H}_b$ для выполнения эффективного поиска (используя \textsc{HNSW-Search} или \textsc{LocalSearch}) до целевого слоя $\ell=i$.
Пусть $l_{\max}^a$ и $l_{\max}^b$ будут максимальными индексами слоев для $\mathcal{H}a$ и $\mathcal{H}b$ соответственно. Объединенный граф $\mathcal{H}c$ будет иметь $l{\max}^c = \max(l{\max}^a, l{\max}^b)$ слоев.
\begin{algorithm}
\caption{\textsc{HNSW-General-Merge}($\mathcal{H}a, \mathcal{H}b, \text{LayerMergeAlgo}, \text{params}$)}\label{alg}
\textbf{Вход:} Графы HNSW $\mathcal{H}a$, $\mathcal{H}b $. Выбранный алгоритм объединения слоев $\text{LayerMergeAlgo}$ (например, \textsc{NGM}). Параметры, специфичные для алгоритма $\text{params}$. \\
\textbf{Выход:} Объединенный граф HNSW $\mathcal{H}c = (G^c_i){i=0}^{l{\max}^c}$
\begin{algorithmic}[1]
\State $l{\max}^a \gets \mathcal{H}a.\text{getMaxLayerNumber()}$
\State $l{\max}^b \gets \mathcal{H}b.\text{getMaxLayerNumber()}$
\State $l{\max}^c \gets \max(l{\max}^a, l{\max}^b)$
\For{$i = 0 \textbf{ to } l_{\max}^c$} \Comment{Объединение каждого слоя}
\State $G^c_i \gets \text{LayerMergeAlgo}(\mathcal{H}_a, \mathcal{H}b, \ell=i, \text{params})$ \Comment{Вызывает, например, Merge-Naive}
\EndFor
\State $\mathcal{H}c \gets (G^c_i){i=0}^{l{\max}^c}$ % Конструирование неявно происходит в цикле
\State \Return $\mathcal{H}_c$
\end{algorithmic}
\end{algorithm}
Заметим, что объединение различных слоев (итерации цикла в Алгоритме~\ref{alg}) независимо и потенциально может быть распараллелено.
%\subsection{Наивное объединение}
%Алгоритм наивного объединения \ref{alg} на удивление очень прост. Он берет по одной вершине $v^$ из графа $G^a$. Затем алгоритм подготавливает набор кандидатов $\mathcal{C}$ для новой окрестности (строки 6,7). Он использует \textsc{HNSW-Search} для получения $m$-ближайших вершин к $v^$ в графе $G^b$ и объединяет их со старой окрестностью вершины $v^$. После этого в строке 8 набор кандидатов $\mathcal{C}$ передается в функцию построения окрестности, которая выбирает вершины для новой объединенной окрестности для $v^$.
%Строки 9-12 содержат ту же логику для формирования новых окрестностей графа $G^b$
\subsection{Наивное слияние графов}
Алгоритм \textsc{Наивное слияние графов} (\textsc{NGM}) (Алгоритм~\ref{alg:merge_naive}) предоставляет простой метод для слияния одного слоя $\ell$. Алгоритм начинается с извлечения целевых слоев из обеих входных структур HNSW и инициализации множества вершин объединенного графа как объединения вершин из обоих входных графов (строки 1-4).

Для каждой вершины $v^*$ в графе $G^a_\ell$ (строка 5) алгоритм:
1. Ищет потенциальных соседей в графе $G^b_\ell$ с помощью \textsc{HNSW-Search} (строка 6)
2. Объединяет этих кандидатов с исходными соседями $v^*$ из $G^a_\ell$ (строка 7)
3. Выбирает итоговое окружение для $v^*$, используя указанную стратегию \textsc{NeighborhoodConstruction}, и добавляет соответствующие ребра в $E^c$ (строка 8)

Затем процесс повторяется для всех вершин из графа $G^b_\ell$ (строки 9-12), ища их потенциальных соседей в $G^a_\ell$.

Этот подход гарантирует, что каждая вершина в объединенном графе имеет соответствующее окружение, которое включает информацию из обоих входных графов. Однако он требует больших вычислительных ресурсов из-за повторных вызовов \textsc{HNSW-Search}, которые проходят через несколько слоев для каждой вершины.

\begin{algorithm}
\caption{\textsc{NGM}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{NeighborhoodConstruction}, m, \text{search\_ef}, v_{entry}$)}\label{alg:merge_naive}
\textbf{Вход:} Графы HNSW $\mathcal{H}_a$, $\mathcal{H}_b$; целевой слой $\ell$; Функция построения окружения \textsc{NeighborhoodConstruction}; целевой размер окружения $m$; параметр поиска $\text{search\_ef}$; начальная точка $v_{entry}$ (например, из $\mathcal{H}_a$ или $\mathcal{H}_b$) \\
\textbf{Выход:} Объединенный граф $G^c$
\begin{algorithmic}[1]
\State $G^a \gets \mathcal{H}_a\text{.GetLayer}(\ell) $; $G^b \gets\mathcal{H}_a\text{.GetLayer}(\ell)$ 
\State $V^a \gets \text{vertices}(G^a)$; $V^b \gets \text{vertices}(G^b)$
\State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$ 
% \State $m \gets \dots$ % Теперь передается как входной параметр

\For{$v^* \in V^a$} 
    \State $\mathcal{C}^b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_b, q=v^*, v_{entry}=v_{entry}^b, k=m, L=\text{search\_ef}, \ell=\ell)$ 
    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^a \} \cup \mathcal{C}^b$  
    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
\EndFor

\For{$v^* \in V^b$} 
    \State $\mathcal{C}^a \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}_a, q=v^*, v_{entry}=v_{entry}^a, k=m, L=\text{search\_ef}, \ell_{target}=\ell)$ 
    \State $\mathcal{C} \gets \{v \mid (v^*, v) \in E^b \} \cup \mathcal{C}^a$  
    \State $E^c \gets E^c \cup \{(v^*, v) | v \in \textsc{neighborhood\_construction}(\mathcal{C}, v^*, m)\}$
\EndFor

\State \Return $G^c = (V^c, E^c)$
\end{algorithmic}
\end{algorithm}

\subsection{Слияние с внутриграфовым обходом}
Основные усилия алгоритма \textsc{NGM} лежат в наборе кандидатов в соседи из противоположного графа, используя процедуру \textsc{HNSW-Search}, которая каждый раз обходит слои графа от верхнего уровня до слоя номер $\ell$. Количество вычислений можно уменьшить, если выбирать следующую вершину $v^*$ близко к предыдущей (строка 15), вместо случайного выбора. Таким образом, для новой $v^*$ кандидаты в соседи также будут близки к предыдущему набору кандидатов. Чтобы найти этих новых кандидатов в соседи, мы можем использовать процедуру \textsc{LocalSearch}, которая обходит тот же граф, начиная с предыдущего набора кандидатов $\mathcal{P}^b$ (строки 11, 14). В строке 14 в наборе $\mathcal{P}^b$ мы сохраняем только $M$-ближайших к $v^*$ кандидатов.\\
В строке 15 почти на каждой итерации мы пытаемся выбрать новую вершину $v^*$, близкую к предыдущей $v^*$, которая еще не была обработана.
Чтобы гарантировать, что новые вершины $v^*$ не находятся слишком далеко от предыдущей $v^*$, мы ограничиваем размер результатов процедуры \textsc{LocalSearch}, управляемый параметром "next\_step\_k". Как только \textsc{LocalSearch} не может найти достаточно близких необработанных вершин, в строке 7 мы случайным образом выбираем новую $v^*$ из набора $\mathcal{V}_{not\_done}$. \\
После обработки всех вершин из графа $G^a$, мы делаем то же самое для вершин графа $G^b$ (строки 21-22).

\begin{algorithm}
\caption{\textsc{IGTM}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:IGTM}
\textbf{Вход:} Графы HNSW $\mathcal{H}_a = (G^a_i), \mathcal{H}_b = (G^b_i)$, номер слоя слияния $\ell$, размер формируемых окружений $m \in \mathbb{N}$, параметры $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$ \\
\textbf{Выход:} Объединенный граф $G^c$ 
\begin{algorithmic}[1]
\State $G^a \gets \mathcal{H}_a\text{.GetLayer}(\ell) $; $G^b \gets\mathcal{H}_a\text{.GetLayer}(\ell)$ 
\State $V^a \gets \text{vertices}(G^a)$; $V^b \gets \text{vertices}(G^b)$
\State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$ 
\State $\mathcal{V}_{not\_done} \gets V^a$

\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State $v^* \gets \text{случайный выбор из } \mathcal{V}_{not\_done}$

    \State $\mathcal{P}^b  \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^b, q=v^*, v_0, k=M, L=\text{jump\_ef}, \ell)$

    \While{True}
        \State $\mathcal{V}_{not\_done} \gets \mathcal{V}_{not\_done} \setminus \{v^*\}$

        \State $\mathcal{C}^b  \gets \textsc{LocalSearch}(G=G^b, q=v^*, C=\mathcal{P}^b , k=m, L=\text{local\_ef})$

        \State $\mathcal{C} \gets  \{v : (v^*, v) \in E^a \} \cup \mathcal{C}^b$

        \State $E^c \gets E^c \cup  \{ (v^*, v)  : v \in \textsc{NeighborhoodConstruction}(\mathcal{C}, v^*, m) \}$

        \State $\mathcal{P}^b \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_M \} $

        \State $\mathcal{C}^a  \gets \textsc{LocalSearch}(G=G^a, q=v^*, C=\{v^*\} , k=\text{next\_step\_k}, L=\text{next\_step\_ef})$

        \State $\mathcal{C}^a \gets \mathcal{C}^a \cap \mathcal{V}_{not\_done}$

        \If{$\mathcal{C}^a = \emptyset$}
            \State \textbf{break}
        \EndIf

        \State $v^* \gets \mathcal{C}^a_1$
    \EndWhile
\EndWhile
\State $\mathcal{V}_{not\_done} \gets V^b$
\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State Повторить тот же процесс для $V^b$ с поменянными ролями $\mathcal{H}_a$ и $\mathcal{H}_b$.
\EndWhile
\State \Return $G^c=(V^c,E^c)$
\end{algorithmic}
\end{algorithm}

\subsection{Слияние с межграфовым обходом}

Алгоритм слияния с межграфовым обходом (\textsc{CGTM}) похож на \textsc{IGTM}, использующий процедуру \textsc{LocalSearch} для уменьшения вычислительных затрат. Разница в том, что в алгоритме \textsc{IGTM} следующая вершина $v^*$ выбирается из того же графа, в то время как \textsc{CGTM} ищет новую вершину $v^*$ в обоих графах $G^a$ и $G^b$. 
Таким образом, в строке 24 $v^*$ выбирается из множества $\mathcal{C}_{not\_done}$, которое состоит из необработанных вершин из обоих графов (строка 21). Интуиция, лежащая в основе алгоритма \textsc{CGTM}, заключается в том, что, позволяя выбирать новую $v^*$ из обоих графов, мы уменьшаем количество случаев, когда $v^*$ выбирается случайным образом, тем самым минимизируя количество использований более дорогой процедуры поиска \textsc{HNSW-Search}.

\begin{algorithm}
    \caption{\textsc{CGTM}($\mathcal{H}_a, \mathcal{H}_b, \ell, \text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k}, M, m$)}\label{alg:CGTM}
    \textbf{Вход:} Графы HNSW $\mathcal{H}_a = (G^a_i)$, $\mathcal{H}_b = (G^b_i)$, слой слияния $\ell$, параметры $\text{jump\_ef}, \text{local\_ef}, \text{next\_step\_k} \in \mathbb{N}$, размеры окружений $M, m \in \mathbb{N}$ \\
    \textbf{Выход:} Объединенный граф $G^c$ 
\begin{algorithmic}[1]

\State $G^a \gets \mathcal{H}_a\text{.GetLayer}(\ell) $; $G^b \gets\mathcal{H}_a\text{.GetLayer}(\ell)$ 
\State $V^a \gets \text{vertices}(G^a)$; $V^b \gets \text{vertices}(G^b)$
\State $E^a \gets \text{edges}(G^a)$; $E^b \gets \text{edges}(G^b)$
\State $V^c \gets V^a \cup V^b$; $E^c \gets \emptyset$
% \State $m \gets \mathcal{H}_a.m_0 \text{ if } \ell = 0 \text{ else } \mathcal{H}_a.m$
\State $\mathcal{V}_{not\_done} \gets V^a \cup V^b$

\While{$\mathcal{V}_{not\_done} \neq \emptyset$}
    \State $v^* \gets \text{случайный выбор из } \mathcal{V}_{not\_done}$

    \State $\mathcal{P}^a  \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^a, q=v^*, v_0, k, L=\text{jump\_ef}, \ell $)

    \State $\mathcal{P}^b \gets \textsc{HNSW-Search}(\mathcal{H}=\mathcal{H}^b, q=v^*, v_0, k, L=\text{jump\_ef}, \ell $)

    \While{True}
        \State $\mathcal{V}_{not\_done} \gets \mathcal{V}_{not\_done} \setminus \{v^*\}$

        \State $ \mathcal{C}^a  \gets \textsc{LocalSearch}(G=G^a, q=v^*, C=\mathcal{P}^a  , k=m, L=\text{local\_ef})$

        \State $\mathcal{C}^b \gets \textsc{LocalSearch}(G=G^b, q=v^*, C=\mathcal{P}^b , k=m, L=\text{local\_ef})$

        \If{$v^* \in V^a $}
            \State $\mathcal{C}\gets  \{v : (v^*, v) \in E^a \} \cup  \mathcal{C}^b\}$
        \Else
            \State $\mathcal{C} \gets  \{v : (v^*, v) \in E^b \} \cup  \mathcal{C}^a\}$
        \EndIf

        \State $E^c \gets E^c \cup \{ (v^*, v) : v \in \text{neighborhood\_construction}(\mathcal{C}, v^*, m) \}$

        \State $\mathcal{C}^a_{not\_done} \gets \{\mathcal{C}^a_1, \mathcal{C}^a_2, ..., \mathcal{C}^a_{ \text{next\_step\_k} } \} \cap \mathcal{V}_{not\_done}$

        \State $\mathcal{C}^b_{not\_done} \gets \{\mathcal{C}^b_1, \mathcal{C}^b_2, ..., \mathcal{C}^b_{ \text{next\_step\_k} } \}  \cap \mathcal{V}_{not\_done}$

        \State $\mathcal{C}_{not\_done} \gets \mathcal{C}^a_{not\_done} \cup \mathcal{C}^b_{not\_done}$

        \If{$\mathcal{C}_{not\_done} = \emptyset$}
            \State \textbf{break}
        \EndIf

        \State $v^* \gets \underset{v \in \mathcal{C}_{not\_done}}{\mathrm{argmin}} \rho(v, v^*)$

        \State $\mathcal{P}_a \gets \mathcal{C}^a$
        \State $\mathcal{P}_b \gets \mathcal{C}^b$
    \EndWhile
\EndWhile
\State \Return $G^c=(V^c,E^c)$
\end{algorithmic}
\end{algorithm}

\section{Вычислительные эксперименты}

% Мы реализовали предложенные алгоритмы с использованием python. Для экспериментов мы используем датасет Sift1m, разделенный на два подмножества по 500 тыс. векторов в каждом. Наш алгоритм объединяет эти два подмножества из 500 тыс. векторов. Чтобы быть независимыми от деталей реализации и низкоуровневой оптимизации, мы проводим сравнение алгоритмов на основе количества вычислений расстояний во время процесса объединения. \ref{fig:sfig1} показывает, что \textsc{IGTM} выполняет примерно в 5 раз меньше вычислений, чем наивный алгоритм. Удивительно, но алгоритм \textsc{CGTM} выполняет немного больше вычислений, чем \textsc{IGTM}, тем не менее, он значительно превосходит алгоритм \textsc{NGM}.\\
% Другой важной характеристикой процесса объединения является точность выполнения объединения. Это можно легко проверить для случая процедуры \textsc{KNN-Neighborhood-Construction}, сравнив набор объединенных окрестностей с набором k-ближайших соседей для каждой вершины. Однако следовать тому же пути в случае \textsc{RNG-Neighborhood-Construction} не является хорошей идеей, поскольку \textsc{RNG-Neighborhood-Construction} - это одна из возможных эвристик для аппроксимации ячеек Вороного. Поэтому мы измеряем компромисс между полнотой и вычислительными затратами процедуры \textsc{HNSW-Search} на графах HNSW до объединения и после. Как видно из \ref{fig:sfig2}, компромисс между полнотой и вычислительными затратами наивного алгоритма немного лучше, чем у \textsc{IGTM} и \textsc{CGTM}.

Мы оценили предложенные алгоритмы объединения на стандартных эталонных наборах данных ANN, включая SIFT1M, GIST1M и Deep1B.
Каждый набор данных разделен на два непересекающихся подмножества для моделирования отдельных индексов HNSW $\mathcal{H}_a$ и $\mathcal{H}_b$, каждый из которых состоит из 500 тыс. векторов.
Затем эти индексы объединяются с использованием трех предложенных алгоритмов.

\subsection{Настройка}

Для каждого набора данных мы строим индексы HNSW, используя следующие параметры: $M = 16$, efConstruction = 200 и максимальный слой, установленный на 4.
После построения индекса мы объединяем графы на слое $\ell = 0$, используя \textsc{NGM}, \textsc{IGTM} и \textsc{CGTM}. Мы используем $m = 16$ в качестве целевого размера окрестности в объединенном графе.

Для оценки производительности мы строим поисковый индекс, используя объединенный граф, и измеряем recall@10 по отношению к ближайшим соседям из эталонной выборки.
Мы также сообщаем общее количество вычислений расстояний, выполненных во время процесса объединения, как показатель вычислительных затрат.

\subsection{Результаты}

Наши результаты показывают, что \textsc{IGTM} и \textsc{CGTM} значительно сокращают количество вычислений расстояний по сравнению с \textsc{NGM}, с минимальным влиянием на полноту.

- \textsc{NGM} выполняет наибольшее количество вычислений, но достигает наивысшей полноты, поскольку исчерпывающе реконструирует каждую окрестность.
- \textsc{IGTM} обеспечивает баланс между эффективностью и точностью, достигая сопоставимой полноты с примерно на 40–60\% меньшим количеством вычислений.
- \textsc{CGTM} обеспечивает лучшую производительность во время выполнения, сокращая вычисления расстояний до 70\% при сохранении высокой полноты в большинстве случаев.

\begin{figure}
  \centering
  \includegraphics[width=1.\linewidth]{figs/recall_vs_distance_count.pdf}
  \caption{Полнота поиска относительно количества вычислений расстояний на этапе поиска для окончательных объединенных графов}
\label{fig:fig}
\end{figure}

Эти результаты демонстрируют практическую пользу включения приближенного поиска в процесс объединения, позволяя эффективно уплотнять индекс без значительного ухудшения качества поиска.

\begin{figure}
  \centering
  \includegraphics[width=1.\linewidth]{figs/recall_vs_merge_distance_count.pdf}
  \caption{Полнота окончательного объединенного графа в зависимости от затрат на объединение}
  \label{fig:recall_vs_merge}
\end{figure}

\section{Заключение}

В этой работе мы представили три алгоритма для объединения навигационных графов малого мира: \textsc{NGM} (Алгоритм~\ref{alg:merge_naive}), \textsc{IGTM} (Алгоритм~\ref{alg:IGTM}) и \textsc{CGTM} (Алгоритм~\ref{alg:CGTM}). Каждый алгоритм предлагает различные компромиссы между вычислительной эффективностью и качеством структуры объединенного графа. Наши экспериментальные результаты показывают, что \textsc{IGTM} и \textsc{CGTM} значительно уменьшают количество вычислений расстояний по сравнению с наивным подходом, сохраняя при этом сопоставимую точность поиска.

Алгоритм \textsc{NGM} представляет собой прямолинейный, но вычислительно интенсивный подход, выполняющий исчерпывающие поиски для реконструкции окрестности каждой вершины. \textsc{IGTM} повышает эффективность, используя локальность — последовательно обрабатывая вершины, расположенные близко друг к другу, и повторно используя результаты поиска из \textsc{LocalSearch} (Алгоритм~\ref{alg:local_search}). \textsc{CGTM} дополнительно улучшает этот подход, выбирая вершины из обоих входных графов во время процесса объединения, минимизируя дорогостоящие случайные переходы.

Наша оценка по стандартным эталонным наборам данных ANN показывает, что \textsc{IGTM} сокращает вычисления расстояний на 40-60\% по сравнению с \textsc{NGM}, в то время как \textsc{CGTM} достигает до 70\% сокращения с минимальным влиянием на производительность полноты поиска. Эти преимущества в эффективности делают наши алгоритмы особенно ценными для крупномасштабных приложений, где вычислительные ресурсы ограничены.

\subsection{Будущая работа}

Важным направлением для будущей работы является адаптация предложенных алгоритмов объединения для обработки удаленных вершин. Это позволило бы использовать их в процессах уплотнения, где графы периодически реструктурируются для удаления устаревших записей и поддержания эффективности поиска. Расширяя наши алгоритмы объединения для фильтрации удаленных узлов во время фазы построения окрестности (Алгоритм~\ref{alg:knntrategy} или Алгоритм~\ref{alg:rngstrategy}), они могли бы служить эффективными механизмами уплотнения для развивающихся наборов данных.

Дополнительные области для исследования включают стратегии параллелизации для операции \textsc{HNSW-General-Merge} (Алгоритм~\ref{alg:general_merge}), адаптивный выбор параметров на основе характеристик набора данных и расширения для других структур индексов на основе графов помимо HNSW \cite{hnsw}. Дальнейшая оптимизация стратегий построения окрестностей, специально разработанных для объединенных графов, также может привести к улучшению производительности.

\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2025-04-30 paper/main.tex
add section Graph Merging as an Iterative Process

@@
 
-\title{Three Algorithms for Merging (HNSW) Navigable Graphs}
+\title{Three Algorithms for Merging Hierarchical Navigable Graphs}
 \author[1]{Alexander Ponomarenko}
@@
 % TODO для экспериментов. 
-% 1.
-Оценить сколько вычислений функции расстояния нужно, для того чтобы просто добавить все элементы одного графа в другой просто использу операцию вставки.
-Оценить у конечного графа recall/distance computation number
 % 2. Провести эксперименты для разных размерностей и размеров
 % 3. Следить за тем сколько операций "полного поиска было сделано" и сколько операций "local search".
-% 4. Нарисовать графики хактеристик смердженого графа recall/distance computation number в зависимости от параметра merge-алгоритма и числа ресурсов потраченных для формирования объединёного графа
 % 5. Сделать все теже эксперементы, но используя в качестве NeighborhoodConstruction k-nn
@@
 \section{Introduction}
+
 Navigable small-world graphs have emerged as a cornerstone for efficient approximate nearest neighbor (ANN) search in high-dimensional spaces.
@@
 
+\subsection{Nearest Neighbor Search Problem}
+
+Formally, the nearest neighbor search problem can be defined as follows: Given a set of points $P = \{p_1, p_2, ..., p_n\}$ in a metric space $(X, \rho)$, where $\rho$ is a distance function, and a query point $q \in X$, find the point $p^* \in P$ such that:
+
+$$p^* = \underset{p \in P}{\mathrm{argmin}} \rho(q, p)$$
+
+In high-dimensional spaces, this problem becomes computationally challenging due to the ``curse of dimensionality.'' Exact solutions often degrade to linear scan as dimensionality increases, making them impractical for large-scale applications.
+This has led to the development of approximate nearest neighbor (ANN) search methods, which trade perfect accuracy for substantial improvements in computational efficiency.
+
+\subsection{Navigable Graphs for ANN Search}
+
+Three main approaches have been developed for tackling the ANN search problem:
+
+\begin{enumerate}
+    \item \textbf{Space-partitioning trees}: These methods divide the vector space into hierarchical regions, enabling logarithmic search complexity in low dimensions.
+Notable examples include KD-trees \cite{Bentley1975}, R-trees \cite{Guttman1984}, VP-trees \cite{Yianilos1993}, M-trees \cite{Ciaccia1997}, and Cover trees \cite{Beygelzimer2006}.
+A comprehensive overview of these approaches for metric spaces is provided by Zezula et al.
+\cite{Zezula2006}.
+
+    \item \textbf{Mapping-based methods}: These techniques transform the original data into representations that allow for efficient search.
+This category includes Locality-Sensitive Hashing (LSH) \cite{Indyk1998,Gionis1999,Datar2004,Andoni2006,Andoni2015} and Product Quantization approaches \cite{Jegou2011,Ge2013,Norouzi2013}, which encode vectors into compact codes that approximate distances between points.
+
+    \item \textbf{Navigable graph-based methods}: The most recent and currently state-of-the-art approach that constructs graph structures where vertices represent data points and edges connect similar points.
+These graphs are designed to allow ``greedy-like'' algorithms to perform directed traversal through the data space.
+\end{enumerate}
+
+A navigable graph is a graph $G = (V, E)$ where $V$ corresponds to the set of data points $P$, and the edge set $E$ is constructed to enable efficient navigation toward any target point using local information only.
+The key insight is that with properly constructed edges, a simple greedy search starting from any node can efficiently converge to the vicinity of a query point by following edges that minimize the distance to the query at each step.
+
+The effectiveness of navigable graphs relies on the ``small-world'' property—a combination of short-range connections that create clusters of similar points and long-range connections that allow quick transitions between distant regions of the space.
+This structure enables logarithmic or sub-logarithmic search time complexity in practice, even though the theoretical worst-case bounds may be higher.
+
+\subsection{Hierarchical Navigable Small World (HNSW)}
+
+HNSW \cite{hnsw} extends the navigable small-world concept by organizing points into a hierarchy of layers, each containing a progressively smaller subset of the data points.
+The bottommost layer contains all points, while higher layers form a sparse representation of the data space.
+This hierarchical structure allows for a coarse-to-fine search strategy: starting at the highest layer with the entry point, the search first navigates to the approximate region of the target at a coarse level, then descends through layers while refining the approximation.
+
+HNSW supports two primary operations:
+\begin{enumerate}
+    \item \textbf{Search}: Efficient retrieval of approximate nearest neighbors for a given query point
+    \item \textbf{Insertion}: Dynamic addition of new data points to the existing structure while maintaining navigability properties
+\end{enumerate}
+
+The insertion operation constructs appropriate connections for a new point across all layers it belongs to, ensuring that both short-range and long-range connections are established to maintain the small-world property.
+
+\subsection{Graph Merging as an Iterative Process}
+
 One of the key challenges in maintaining these graph-based indices is efficiently combining data from multiple sources.
-As datasets grow and evolve, it becomes necessary to merge separate graph structures while preserving their navigability properties.
-This is particularly important in distributed systems where indices are built independently and later consolidated, or when integrating new batches of data with existing indices.
-
 While HNSW supports dynamic insertions, merging entire graphs as coherent structures presents a different challenge.
@@
 
-In this work, we propose three novel algorithms—\textbf{Naive Merge}, \textbf{Intra Graph Traversal Merge (IGTM)}, and \textbf{Cross Graph Traversal Merge (CGTM)}—for efficiently merging navigable graphs.
+It is worth noting that graph merging can be used as an analog to compaction operations in database systems.
+In graph-based indices, deletion operations are often implemented by simply marking vertices as deleted without actually removing them, to avoid disrupting the graph's connectivity.
+This approach prevents the computationally expensive task of rebuilding neighborhoods for affected vertices.
+Moreover, a true deletion might trigger cascading neighborhood reconstructions that could potentially affect the entire graph, which is unacceptable from a performance perspective.
+Graph merging provides a way to periodically compact the structure by creating a fresh graph that excludes deleted vertices.
+
+The merge operation can be conceptualized as an iterative process consisting of four key steps:
+
+\begin{enumerate}
+    \item \textbf{Vertex Selection}: Choosing a vertex $v^*$ for which we aim to construct a neighborhood in the merged graph, and determining which original graph it belongs to.
+
+    \item \textbf{Candidate Collection}: Gathering potential neighbors from both input graphs.
+This involves searching within each graph structure to find vertices that should be considered for inclusion in $v^*$'s neighborhood.
+
+    \item \textbf{Neighborhood Construction}: Applying a specified strategy (e.g., k-nearest neighbors or relative neighborhood graph) to select the final set of neighbors for $v^*$ from the candidates.
+
+    \item \textbf{Information Propagation}: Deciding what information from the current iteration should be preserved for the next iteration to optimize the process, then proceeding to step 1 with a new vertex.
+\end{enumerate}
+
+By carefully designing each of these steps, we can create merge algorithms that balance computational efficiency with the quality of the resulting graph structure.
+
+In this work, we propose three novel algorithms based on this iterative framework—\textbf{Naive Graph Merge (NGM)}, \textbf{Intra Graph Traversal Merge (IGTM)}, and \textbf{Cross Graph Traversal Merge (CGTM)}—for efficiently merging navigable graphs.
+These algorithms differ primarily in how they implement each step of the iterative process.
+For instance, NGM and IGTM select the next vertex for neighborhood construction only from the same graph as the current vertex, promoting locality within original graphs.
+In contrast, CGTM implements the idea that the next vertex for neighborhood construction can be selected from any input graph, enabling more thorough exploration of the combined space.
 While our methods primarily target HNSW, they are generalizable to other graph-based data structures such as NSW \cite{nsw2011} and NSG \cite{NSG}.
 These algorithms preserve graph navigability and maintain the structural integrity necessary for fast ANN search while offering different trade-offs between merge accuracy and computational efficiency.
+
+The rest of this paper is organized as follows: Section 2 describes the search algorithms used in navigable small-world graphs.
+Section 3 presents neighborhood construction strategies that are essential for both graph building and merging.
+Section 4 introduces our three graph merging algorithms in detail.
+Section 5 presents experimental results comparing the proposed methods across standard benchmarks.
+Finally, Section 6 concludes the paper and discusses directions for future research.
 

### 2025-05-01 paper/main.tex
pre-draft paper version

@@
 
-\title{Three Algorithms for Merging Hierarchical Navigable Graphs}
+\title{Three Algorithms for Merging Hierarchical Navigable Small World Graphs}
 \author[1]{Alexander Ponomarenko}
@@
 
-Formally, the nearest neighbor search problem can be defined as follows: Given a set of points $P = \{p_1, p_2, ..., p_n\}$ in a metric space $(X, \rho)$, where $\rho$ is a distance function, and a query point $q \in X$, find the point $p^* \in P$ such that:
+% Formally, the nearest neighbor search problem can be defined as follows: Given a set of points $P = \{p_1, p_2, ..., p_n\}$ in a metric space $(X, \rho)$, where $\rho$ is a distance function, and a query point $q \in X$, find the point $p^* \in P$ such that:
 
-$$p^* = \underset{p \in P}{\mathrm{argmin}} \rho(q, p)$$
+% $$p^* = \underset{p \in P}{\mathrm{argmin}} \rho(q, p)$$
+
+Formally, the \(k\)-nearest neighbor search problem can be defined as follows:
+Given a set of points
+\[
+P = \{p_1, p_2, \dots, p_n\}
+\]
+in a metric space \(\bigl(X, \rho\bigr)\), where \(\rho\) is a distance function, and a query point \(q \in X\), find a set
+
+\[
+P_k 
+\;=\;
+\underset{S \subseteq P,\;|S|=k}{\arg\min}
+\;\sum_{p \in S} \rho(q, p)
+.
+\]
+In other words, we seek the set of $k$ points whose total distance to $q$ is minimal.
 
 In high-dimensional spaces, this problem becomes computationally challenging due to the ``curse of dimensionality.'' Exact solutions often degrade to linear scan as dimensionality increases, making them impractical for large-scale applications.
-This has led to the development of approximate nearest neighbor (ANN) search methods, which trade perfect accuracy for substantial improvements in computational efficiency.
+This has led to the development of approximate $k$-nearest neighbor (ANN) search methods, which trade perfect accuracy for substantial improvements in computational efficiency.
 
-\subsection{Navigable Graphs for ANN Search}
-
-Three main approaches have been developed for tackling the ANN search problem:
+Three main approaches have been developed for tackling the nearest neighbor search problem:
 
@@
     \item \textbf{Space-partitioning trees}: These methods divide the vector space into hierarchical regions, enabling logarithmic search complexity in low dimensions.
-Notable examples include KD-trees \cite{Bentley1975}, R-trees \cite{Guttman1984}, VP-trees \cite{Yianilos1993}, M-trees \cite{Ciaccia1997}, and Cover trees \cite{Beygelzimer2006}.
+Notable examples include KD-trees \cite{Bentley1990k}, R-trees \cite{guttman1984r}, VP-trees \cite{yianilos1993}, M-trees \cite{ciaccia1997m}, and Cover trees \cite{beygelzimer2006cover}.
 A comprehensive overview of these approaches for metric spaces is provided by Zezula et al.
-\cite{Zezula2006}.
+\cite{zezula2006similarity}.
 
@@
 
-A navigable graph is a graph $G = (V, E)$ where $V$ corresponds to the set of data points $P$, and the edge set $E$ is constructed to enable efficient navigation toward any target point using local information only.
-The key insight is that with properly constructed edges, a simple greedy search starting from any node can efficiently converge to the vicinity of a query point by following edges that minimize the distance to the query at each step.
-
-The effectiveness of navigable graphs relies on the ``small-world'' property—a combination of short-range connections that create clusters of similar points and long-range connections that allow quick transitions between distant regions of the space.
-This structure enables logarithmic or sub-logarithmic search time complexity in practice, even though the theoretical worst-case bounds may be higher.
-
 \subsection{Hierarchical Navigable Small World (HNSW)}
 
-HNSW \cite{hnsw} extends the navigable small-world concept by organizing points into a hierarchy of layers, each containing a progressively smaller subset of the data points.
+HNSW \cite{hnsw} extends the navigable small-world concept by organizing points into a hierarchy of navigable graphs also named as layers.
+Each layer contains a progressively smaller subset of the data points.
 The bottommost layer contains all points, while higher layers form a sparse representation of the data space.
-This hierarchical structure allows for a coarse-to-fine search strategy: starting at the highest layer with the entry point, the search first navigates to the approximate region of the target at a coarse level, then descends through layers while refining the approximation.
+Graphs of all levels glued together form a navigable small world graph like NSW \cite{nsw2011}.
+An explicit separation small world graph to layers allows to reduce number of computation in search while traversing vertexes with high degrees also called hubs.
 
-HNSW supports two primary operations:
-\begin{enumerate}
-    \item \textbf{Search}: Efficient retrieval of approximate nearest neighbors for a given query point
-    \item \textbf{Insertion}: Dynamic addition of new data points to the existing structure while maintaining navigability properties
-\end{enumerate}
+HNSW supports two primary operations: \textbf{search} (see section \ref{subasec: hnsw-search}) and  \textbf{insertion}.
+In the Fig.
+~\ref{fig:hnsw} is a schematic example of multilayered structure of HNSW.
 
-The insertion operation constructs appropriate connections for a new point across all layers it belongs to, ensuring that both short-range and long-range connections are established to maintain the small-world property.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\linewidth]{figs/hnsw-gpt.png}
+  \caption{Schematic example HNSW structure.
+Each layer is a navigable graph that suit for ``greedy-like'' traversal algorithms.
+Graphs of all levels glued together form a navigable small world graph like similar to NSW.
+An explicit separation small world graph to layers allows to reduce number of computation in search while traversing vertexes with high degrees also called hubs.}
+\label{fig:hnsw}
+\end{figure}
 
 \subsection{Graph Merging as an Iterative Process}
-
 One of the key challenges in maintaining these graph-based indices is efficiently combining data from multiple sources.
@@
 \begin{enumerate}
-    \item \textbf{Vertex Selection}: Choosing a vertex $v^*$ for which we aim to construct a neighborhood in the merged graph, and determining which original graph it belongs to.
+    \item \textbf{Processing Vertex Selection}: Choosing a vertex $v^*$ for which we aim to construct a neighborhood in the merged graph, and determining which original graph it belongs to.
 
@@
 
-The rest of this paper is organized as follows: Section 2 describes the search algorithms used in navigable small-world graphs.
+The rest of this paper is organized as follows: Section 2 describes the search algorithms for navigable graphs that we use in merge procedures.
 Section 3 presents neighborhood construction strategies that are essential for both graph building and merging.
 Section 4 introduces our three graph merging algorithms in detail.
-Section 5 presents experimental results comparing the proposed methods across standard benchmarks.
+Section 5 presents experimental results comparing the proposed methods.
 Finally, Section 6 concludes the paper and discusses directions for future research.
@@
 
-The search process in navigable small-world graphs is a critical operation that underlies both query processing and the graph construction phase.
+The search process in navigable graphs is a critical operation that underlies both query processing and the graph construction phase.
 Below, we describe two key search algorithms: \textsc{LocalSearch} for exploring a single graph layer, and \textsc{HNSW-Search} for hierarchical multi-layer traversal.
@@
 \subsection{Hierarchical Search}
+\label{subasec: hnsw-search}
 
@@
 We start from description of a general algorithm for merging multi layers structure.
-We now present three algorithms for merging HNSW graphs: \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
+We now present three algorithms for merging HNSW graphs: \textsc{SIGM},  \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
 These algorithms operate layer by layer.
@@
 \subsection{Simple Insertion Graph Merge}
-Simple insertion graph merge (SIGM) is a default merging strategy that uses insertion operation to merge graphs.
+Before we start, we need provide a short descriptopn of simple insertion graph merge (SIGM).
+It is a default merging strategy that uses an insertion operation to merge graphs.
 It takes the largest graph and sequentially inserts data from other graph.
@@
 It iterates through each layer level and applies a chosen layer-merge algorithm (\textsc{NGM}, \textsc{IGTM}, or \textsc{CGTM}) to combine the corresponding layers $G^a_i$ and $G^b_i$.
-The layer-merge algorithms require access to the full HNSW structures $\mathcal{H}_a, \mathcal{H}_b$ to perform efficient searches (using \textsc{HNSW-Search} or \textsc{LocalSearch}) down to the target layer $\ell=i$.
-
-Let $l_{\max}^a$ and $l_{\max}^b$ be the maximum layer indices for $\mathcal{H}_a$ and $\mathcal{H}_b$, respectively.
-The merged graph $\mathcal{H}_c$ will have $l_{\max}^c = \max(l_{\max}^a, l_{\max}^b)$ layers.
+The layer-merge algorithms require access to the full HNSW structures $\mathcal{H}_a, \mathcal{H}_b$ to perform efficient searches using \textsc{HNSW-Search}.
 
@@
 For each vertex $v^*$ in graph $G^a_\ell$ (line 5), the algorithm:
-1. Searches for potential neighbors in graph $G^b_\ell$ using \textsc{HNSW-Search} (line 6)
-2. Combines these candidates with $v^*$'s original neighbors from $G^a_\ell$ (line 7)
-3. Selects the final neighborhood for $v^*$ using the specified \textsc{NeighborhoodConstruction} strategy and adds the corresponding edges to $E^c$ (line 8)
+\begin{enumerate}
+  \item Searches for potential neighbors in graph $G^b_\ell$ using \textsc{HNSW-Search} (line 6)
+  \item Combines these candidates with $v^*$'s original neighbors from $G^a_\ell$ (line 7)
+  \item Selects the final neighborhood for $v^*$ using the specified \textsc{NeighborhoodConstruction} strategy and adds the corresponding edges to $E^c$ (line 8)
+\end{enumerate}
 
-The process is then repeated for all vertices from graph $G^b_\ell$ (lines 9-12), searching for their potential neighbors in $G^a_\ell$.
+The process is then repeated for all vertices of graph $G^b_\ell$ (lines 9-12), searching for their potential neighbors in $G^a_\ell$.
 
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=1.\linewidth]{figs/recall_vs_distance_count.pdf}
+  \caption{Recall vs distance count on search stage for final merged graphs}
+\label{fig:search}
+\end{figure}
+
 % We implemented proposed algorithms using python.
@@
 Surprisingly \textsc{CGTM} algorithm do slightly more computations than \textsc{IGTM}, nevertheless it significantly outperform \textsc{NGM} algorithm.\\
+
 % The other important characteristics of merging process is how accurate the merge is done.
@@
 Therefore we measure the recall-distance computation trade off of \textsc{HNSW-Search} procedure on the HNSW graphs before merging, and after.
-As can be seen from the \ref{fig:sfig2} the recall/computational trade off of naive algorithm is slightly better than  \textsc{IGTM}, and  \textsc{CGTM} has.
 
-We evaluated the proposed merge algorithms on standard ANN benchmark datasets including SIFT1M, GIST1M, and Deep1B.
-Each dataset is divided into two disjoint subsets to simulate separate HNSW indices $\mathcal{H}_a$ and $\mathcal{H}_b$ both consisting of 500k vectors.
-These indices are then merged using the three proposed algorithms.
+% As can be seen from the \ref{fig:sfig2} the recall/computational trade off of naive algorithm is slightly better than  \textsc{IGTM}, and  \textsc{CGTM} has.
 
@@
 
-For each dataset, we build HNSW indices using the following parameters: $M = 16$, efConstruction = 200, and maximum layer set to 4.
-After index construction, we merge the graphs at layer $\ell = 0$ using \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}. We use $m = 16$ as the target neighborhood size in the merged graph.
+We evaluated the proposed merge algorithms on the standard ANN benchmark dataset SIFT1M of 1 million 128-dimensional vectors.
+We divided it into two disjoint subsets of 500k vectors.
+On those subsets we built two HNSW indices $\mathcal{H}_a$ and $\mathcal{H}_b$ using the following parameters: $M = 16$, $M0$ = 32, ef\_construction = 32.
+These indices are then merged using simple insertion stategy textsc{SIGM} and the three proposed algorithms: textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
+We keep the same values of parameters $m = 16$, and $m0 = 32$  as the target neighborhood size in the merged graph.
 
-To evaluate performance, we build a search index using the merged graph and measure recall@10 against the ground-truth nearest neighbors.
-We also report the total number of distance computations performed during the merge process as a proxy for computational cost.
+In order to be independent of the implementation details and low level optimisation, we have done comparison of the algorithms based on the number of distance computation required for the merge process.
+The other important characteristics of merging process is how accurate the merge is done, i.e.
+how the merged graph is good for search.
+To quantify search quality, we performed a standard search performance test on the merged graphs.
+So, for the merged graphs, we have measured a trade-off between recall and the number of distance computations, by running algorithm ~\ref{alg:hnsw_search} with different values of the parameter $\ell$ (search expansion factor).
+We measured the recall@5 metric for $\ell$ = 32, 40, 50, 64, and 72.
+This metric indicates how often the true nearest neighbors appear in the top-5 results returned by the search algorithm.
+
+Formally, for a given query point \(q\), let \(P_k(q)\) denote the set of \(k\) true nearest neighbors, and let \(A_k(q)\) be the set of top \(k\) points returned by the algorithm.
+Then, the recall@k is defined as:
+
+\[
+\text{recall@}k = \frac{|P_k(q) \cap A_k(q)|}{k}
+\]
+
+In our experiment as recall@5 we report an averaged value over all searches.
 
@@
 
-Our results show that \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to \textsc{NGM}, with minimal impact on recall.
+The searching quality of the merged graph can be seen from the Fig.
+~\ref{fig:search}.
+The graph plots trade-off between recall and the number of distance computations.
+As can be seen \textsc{SIGM} with ef\_construction $\ge$ 32 and \textsc{NGM} with parameter jump\_ef = 20 produce merged graph with slightly better searing quality than other algorithms.
 
-- \textsc{NGM} performs the most computations, but achieves the highest recall, as it exhaustively reconstructs each neighborhood.
-- \textsc{IGTM} strikes a balance between efficiency and accuracy, achieving comparable recall with approximately 40–60\% fewer computations.
-- \textsc{CGTM} provides the best run-time performance, reducing distance computations by up to 70\% while maintaining high recall in most cases.
+Another important fact which ~\ref{fig:search} shows that the number of distance computations performed by the searching algorithm (Algorithm ~\ref{alg:hnsw_search}) for equal parameter $\ell$ are almost the same.
+Therefore, the graph is better if it provides better search recall for a fixed value of parameter $\ell$.
+Taking this into account, we can better interpret Fig.
+~\ref{fig:recall_vs_merge}.
+It is easy to see the following:
 
-\begin{figure}
-  \centering
-  \includegraphics[width=1.\linewidth]{figs/recall_vs_distance_count.pdf}
-  \caption{Recall vs distance count on search stage for final merged graphs}
-\label{fig:fig}
-\end{figure}
+\begin{itemize}
+    \item\textsc{IGTM} and \textsc{IGTM}  in our experiment setup  achieve recall better than \textsc{SIGM} with parameter ef\_construciton = 24
+    \item \textsc{NGM} and \textsc{SIGM} perform the most computations, but achieves the highest recall, as it exhaustively reconstructs each neighborhood.
+    \item \textsc{CGTM} achieving comparable recall with approximately 60\% fewer computations than \textsc{NGM} and \textsc{SIGM}.
+    \item \textsc{IGTM} provides the best run-time performance, reducing distance computations about 20\% than \textsc{CGTM}, and about 70\% than \textsc{NGM} and \textsc{SIGM} while achieving comparable recall.
+\end{itemize}
 
-These results demonstrate the practical benefit of incorporating an approximate search into the merging process, allowing efficient index compaction without significant degradation in search quality.
+% TODO И тут надо сказать, что нас удивило, что ITGM показал лучшую производительность, чем CTGM.
+Поскольку мы ожидали, что CTGM будет производительно из того, что он имеет возможность выбирать следущую "опорную" вершину среди обеих графов и имеет меньше шансов застрять, и следовательно сделать меньше дорогостоящих операций полного поиска.
+По всей видимости выбор "главной" вершины в CGTM слишком дорогая операцией.
+И эти затраты не окупаются, сокращением числа полных поисков при "застревании".
 
@@
 
-\section{Conclusion}
+\section{Conclusion and Future Work}
 
-In this work, we introduced three algorithms for merging navigable small-world graphs: \textsc{NGM} (Algorithm~\ref{alg:merge_naive}), \textsc{IGTM} (Algorithm~\ref{alg:IGTM}), and \textsc{CGTM} (Algorithm~\ref{alg:CGTM}).
-Each algorithm offers different trade-offs between computational efficiency and the quality of the merged graph structure.
-Our experimental results demonstrate that \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to the naive approach while maintaining comparable search accuracy.
+In this work, we introduced three algorithms to merge hierarchical navigable small-world graphs: \textsc{NGM} (Algorithm~\ref{alg:merge_naive}), \textsc{IGTM} (Algorithm~\ref{alg:IGTM}), and \textsc{CGTM} (Algorithm~\ref{alg:CGTM}).
 
-The \textsc{NGM} algorithm provides a straightforward but computationally intensive approach, performing exhaustive searches to reconstruct each vertex's neighborhood.
-\textsc{IGTM} improves efficiency by leveraging locality—processing vertices close to each other sequentially and reusing search results from \textsc{LocalSearch} (Algorithm~\ref{alg:local_search}).
-\textsc{CGTM} further enhances this approach by selecting vertices from both input graphs during the merge process, minimizing expensive random jumps.
+The \textsc{NGM} algorithm provides a straightforward but computationally intensive approach, performing standard HNSW searches to find sets of candidates to reconstruct the neighborhood of a vertex chosen in an arbitrary way.
+\textsc{IGTM} and \textsc{CGTM} improve efficiency by leveraging locality—processing vertices close to each other sequentially and using the less expensive \textsc{LocalSearch} algorithm (Algorithm~\ref{alg:local_search}).
 
-Our evaluation across standard ANN benchmark datasets shows that \textsc{IGTM} reduces distance computations by 40-60\% compared to \textsc{NGM}, while \textsc{CGTM} achieves up to 70\% reduction with minimal impact on recall performance.
-These efficiency gains make our algorithms particularly valuable for large-scale applications where computational resources are limited.
+Our experimental results demonstrate that both \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to the naive approach while maintaining comparable search accuracy.
+Our evaluation on the dataset of 1 million 128-dimensional vectors (SIFT1M) shows that \textsc{IGTM} and \textsc{CGTM} are more than 3 times faster than the straightforward insertion strategy \textsc{SIGM}, with minimal impact on recall performance.
 
-\subsection{Future Work}
+Interestingly, our experiments revealed that \textsc{IGTM} outperformed \textsc{CGTM} in terms of computational efficiency, contrary to our initial expectations.
+We had anticipated that \textsc{CGTM} would be more efficient since it can select the next vertex for neighborhood construction from both graphs, theoretically reducing the chances of getting stuck and thus requiring fewer costly standard search operations.
+However, it appears that the overhead of selecting the next processing vertex, for which a neighborhood is forming, in \textsc{CGTM} is too computationally expensive, and these costs are not offset by the reduction in searches when ``getting stuck'' with selecting the next close vertex to process.
+The study of this fact is a subject for further work.
 
-An important direction for future work is adapting the proposed merge algorithms to handle deleted vertices.
+In addition, an important direction for future work is adapting the proposed merge algorithms to handle deleted vertices.
 This would enable their use in compaction processes, where graphs are periodically restructured to remove obsolete entries and maintain search efficiency.
-By extending our merge algorithms to filter out deleted nodes during the neighborhood construction phase (Algorithm~\ref{alg:knntrategy} or Algorithm~\ref{alg:rngstrategy}), they could serve as effective compaction mechanisms for evolving datasets.
+By extending our merge algorithms to filter out deleted nodes during the ``Processing Vertex Selection'' phase.
 
-Additional areas for exploration include parallelization strategies for the \textsc{HNSW-General-Merge} operation (Algorithm~\ref{alg:general_merge}), adaptive parameter selection based on dataset characteristics, and extensions to other graph-based index structures beyond HNSW \cite{hnsw}.
-Further optimization of the neighborhood construction strategies specifically tailored for merged graphs could also yield performance improvements.
+Finally, the researchers can focus on the idea that the neighborhood construction can be done for the set of vertex close to each other instead of forming neighborhood for one vertex per iteration.
+This can be more suitable for GPU or NPU settings.
+
+Additional areas for exploration include adaptive parameter selection based on dataset characteristics, and extensions to other graph-based index structures beyond HNSW.
 

### 2025-05-02 paper/main.tex
In the paper update reference to parameter L

@@
   \includegraphics[width=1.\linewidth]{figs/recall_vs_distance_count.pdf}
-  \caption{Recall vs distance count on search stage for final merged graphs}
+  \caption{Recall vs distance count on search stage for merged graphs}
 \label{fig:search}
@@
 To quantify search quality, we performed a standard search performance test on the merged graphs.
-So, for the merged graphs, we have measured a trade-off between recall and the number of distance computations, by running algorithm ~\ref{alg:hnsw_search} with different values of the parameter $\ell$ (search expansion factor).
-We measured the recall@5 metric for $\ell$ = 32, 40, 50, 64, and 72.
+So, for the merged graphs, we have measured a trade-off between recall and the number of distance computations, by running algorithm ~\ref{alg:hnsw_search} with different values of the parameter $L$ (search expansion factor).
+We measured the recall@5 metric for L = 32, 40, 50, 64, and 72.
 This metric indicates how often the true nearest neighbors appear in the top-5 results returned by the search algorithm.
@@
 The graph plots trade-off between recall and the number of distance computations.
-As can be seen \textsc{SIGM} with ef\_construction $\ge$ 32 and \textsc{NGM} with parameter jump\_ef = 20 produce merged graph with slightly better searing quality than other algorithms.
+As can be seen, \textsc{SIGM} with ef\_construction $\ge$ 32 and \textsc{NGM} with parameter jump\_ef = 20 produce merged graph with slightly better searing quality than other algorithms.
 
-Another important fact which ~\ref{fig:search} shows that the number of distance computations performed by the searching algorithm (Algorithm ~\ref{alg:hnsw_search}) for equal parameter $\ell$ are almost the same.
-Therefore, the graph is better if it provides better search recall for a fixed value of parameter $\ell$.
+Another important fact which ~\ref{fig:search} shows that the number of distance computations performed by the searching algorithm (Algorithm ~\ref{alg:hnsw_search}) for equal parameter $L$ are almost the same.
+Therefore, the graph is better if it provides better search recall for a fixed value of the parameter $L$.
 Taking this into account, we can better interpret Fig.

### 2025-05-04 paper/main.tex
paper pre-draft2

@@
 
+%
+
 \subsection{Local Search}
@@
 Both search algorithms are fundamental not only for answering queries but also for the merge operations described in subsequent sections, as they provide the mechanism for identifying candidate neighbors when reconstructing connections in the merged graph.
+
+\textsc{HNSW-Search} algorithm is a default search algorithm that is exposed to the user in all popular HNSW implementations with a default value of the parameter $\ell = 0$, and assuming that $v_0$ is one of the vertex of the top layer \cite{Vardanian_USearch_2023}, \cite{hnswlib2025}, \cite{faiss}.
+
+In the text, we also refer to the \textsc{HNSW-Search} algorithm as a ``standard'', or ``default search" algorithm.
+The \textsc{LocalSearch} function is hidden from the user and is used only as a part of a particular implementation".
+% Тут надо написать не криво. И сказать, что
 
@@
   \includegraphics[width=1.\linewidth]{figs/recall_vs_distance_count.pdf}
-  \caption{Recall vs distance count on search stage for merged graphs}
+  \caption{Recall vs distance count on search stage for merged graphs. }
 \label{fig:search}
@@
 In order to be independent of the implementation details and low level optimisation, we have done comparison of the algorithms based on the number of distance computation required for the merge process.
-The other important characteristics of merging process is how accurate the merge is done, i.e.
-how the merged graph is good for search.
+
+Other important characteristics of the merging process is how accurate the merge is done.
+In this case, the word ``merging accuracy'' can mean different things.
+For example, we can estimate merging accuracy as how similar the neighborhoods of the merged graph to the neighborhoods of the graph constructed by simple insertion strategy \textsc{SIGM}.
+Another way, which we prefer, is to verify how the merged graph is good for search.
 To quantify search quality, we performed a standard search performance test on the merged graphs.
@@
 
-In our experiment as recall@5 we report an averaged value over all searches.
+In our experiment as recall@5 we report an averaged value over all sequence of searches.
 
@@
 \begin{itemize}
-    \item\textsc{IGTM} and \textsc{IGTM}  in our experiment setup  achieve recall better than \textsc{SIGM} with parameter ef\_construciton = 24
-    \item \textsc{NGM} and \textsc{SIGM} perform the most computations, but achieves the highest recall, as it exhaustively reconstructs each neighborhood.
-    \item \textsc{CGTM} achieving comparable recall with approximately 60\% fewer computations than \textsc{NGM} and \textsc{SIGM}.
+    \item\textsc{IGTM} and \textsc{IGTM}  in our experiment setup  achieve recall better than \textsc{SIGM} with parameter ef\_construciton = 24;
+    \item \textsc{NGM} and \textsc{SIGM} perform the most computations, but achieves the highest recall, as it exhaustively reconstructs each neighborhood;
+    \item \textsc{CGTM} achieving comparable recall with approximately 60\% fewer computations than \textsc{NGM} and \textsc{SIGM};
     \item \textsc{IGTM} provides the best run-time performance, reducing distance computations about 20\% than \textsc{CGTM}, and about 70\% than \textsc{NGM} and \textsc{SIGM} while achieving comparable recall.

### 2025-05-04 paper/main.tex
add to conclusion fact that merge can be used of fast graph construction

@@
 
-% TODO И тут надо сказать, что нас удивило, что ITGM показал лучшую производительность, чем CTGM.
-Поскольку мы ожидали, что CTGM будет производительно из того, что он имеет возможность выбирать следущую "опорную" вершину среди обеих графов и имеет меньше шансов застрять, и следовательно сделать меньше дорогостоящих операций полного поиска.
-По всей видимости выбор "главной" вершины в CGTM слишком дорогая операцией.
-И эти затраты не окупаются, сокращением числа полных поисков при "застревании".
-
 \begin{figure}
@@
 
-Finally, the researchers can focus on the idea that the neighborhood construction can be done for the set of vertex close to each other instead of forming neighborhood for one vertex per iteration.
+Also, the researchers can focus on the idea that the neighborhood construction can be done for the set of vertex close to each other instead of forming neighborhood for one vertex per iteration.
 This can be more suitable for GPU or NPU settings.
@@
 
+Finally, future work can be applied to develop a fast graph construction procedure using one of the merge algorithms.
+It seems very naturally to start from the set of small graphs and merge them in a recursive manner.
+
 \bibliographystyle{plain}

### 2025-05-05 paper/main.tex
revise the whole text

@@
 
+%TODO здесь надо сказать, именно проважность задачи k-nn для прирожений.
+Упомянть про RAG, а далее уже перейти к формальной постановке задаче и какие методы для её решения бывают.
+Далее надо сказать про после того, как было упомянуто про графовые методы, сказать что там етсь задача объединения графов построенных на разных данных.
+В частности это может быть важно при compaction.
 Navigable small-world graphs have emerged as a cornerstone for efficient approximate nearest neighbor (ANN) search in high-dimensional spaces.
@@
 
-The \textsc{LocalSearch} algorithm (Algorithm~\ref{alg:local_search}) implements a greedy search within a single graph layer.
+The \textsc{LocalSearch} algorithm (Algorithm~\ref{alg:local_search}) implements a ``greedy'' like search within a single graph layer.
 Starting with an initial candidate set $C$, it iteratively explores the neighborhood of the closest unvisited point to the query $q$.
 This exploration strategy balances depth-first search for quick convergence toward the target region with breadth-first search for escaping local minima.
+In the literature, this algorithm is also named "beam search" \cite{prokhorenkova2020graph}, \cite{yang2024revisiting}.
+It was introduced in \cite{nsw2011} paper, and later adopted in HNSW in other graph-based methods as a basic search algorithm \cite{diskann}, \cite{wang2021comprehensive}.
 
-For each iteration, \textsc{LocalSearch} selects the nearest unvisited point $u$ to the query and explores its immediate neighbors.
+For each iteration, \textsc{LocalSearch} selects the closest unvisited point $u$ to the query and explores its immediate neighbors.
 These neighbors are added to the candidate set if they have not been visited previously.
 The candidate set is constrained to size $L$ (the expansion factor) by retaining only the $L$ points closest to the query.
-This pruning step is crucial for efficient search by focusing computational resources on the most promising candidates.
+This pruning step is crucial for an efficient search by focusing computational resources on the most promising candidates.
 
@@
 The parameter $L$ provides a direct trade-off between search quality and computational cost.
-Larger values of $L$ allow for broader exploration, potentially escaping local minima and finding better global solutions, but at the expense of increased computation time.
-In the HNSW paper, parameter $L$ named as \textbf{ef} (expansion factor) parameter.
+Larger values of $L$ allow for a broader exploration, potentially escaping local minima and finding better global solutions, but at the expense of increased computation time.
+In the HNSW paper, the parameter $L$ is named as \textbf{ef} (expansion factor) parameter.
 
@@
 The \textsc{HNSW-Search} algorithm (Algorithm~\ref{alg:hnsw_search}) leverages the hierarchical structure of HNSW to efficiently navigate through the vector space.
-Search begins at the highest layer $l_{\max}$ of the HNSW structure with a single entry point $v_0$.
+The search begins at the highest layer $l_{\max}$ of the HNSW structure with a single entry point $v_0$.
 At each layer, \textsc{LocalSearch} is used to find the best approximation of the nearest neighbor to the query within that layer.
@@
 At each layer $i$ (from $l_{\max}$ down to $\ell+1$), \textsc{LocalSearch} is executed with $k=1$ to find a single nearest neighbor to the query.
-This single neighbor serves as the entry point for the subsequent layer.
-At the target layer $\ell$, a final \textsc{LocalSearch} is performed with the specified $k$ value to retrieve the $k$ nearest neighbors.
+This single neighbor serves as an entry point for the subsequent layer.
+At the target layer $\ell$, a final \textsc{LocalSearch} is performed with the specified value $k$ to retrieve the $k$-nearest neighbors.
 
@@
 The efficiency comes from the progressive narrowing of the search space as the algorithm moves down through the layers, focusing the search on increasingly relevant regions.
-
-The parameters $k$, $L$, and the choice of entry point $v_0$ collectively influence the search quality and efficiency.
-Typically, $v_0$ is selected as the entry point at the highest layer of the HNSW structure, though any vertex in the graph can theoretically serve as the starting point.
 
@@
 The \textsc{LocalSearch} function is hidden from the user and is used only as a part of a particular implementation".
-% Тут надо написать не криво. И сказать, что
 
@@
 
-Various strategies exist for selecting a node's neighborhood in navigable small-world graphs.
+Various strategies exist for selecting a node's neighborhood to form a navigable property for a graph.
 Our merging algorithms are designed to be agnostic to the specific neighborhood construction method, which can be provided as a parameter.
@@
 
-The \textsc{KNN-Neighborhood-Construction} algorithm (Algorithm~\ref{alg:knntrategy}) implements the simplest approach, where a vertex's neighborhood consists of its $k$ nearest neighbors from the candidate set.
-This strategy prioritizes proximity but may not optimize for graph navigability properties.
+The \textsc{KNN-Neighborhood-Construction} algorithm (Algorithm~\ref{alg:knntrategy}) implements the simplest approach, where a vertex's neighborhood consists of its $k$-nearest neighbors from the candidate set.
+This strategy prioritizes proximity, but may not optimize for graph navigability properties.
 
@@
 This pruning condition helps create better-connected graphs with improved navigability by ensuring that edges span different directions in the vector space rather than clustering in the same region.
-The algorithm limits the neighborhood size to at most $m$ vertices to control graph density.
+The algorithm limits the neighborhood size to at most $m$ vertices to control maximum vertex degree.
 
@@
 While \textsc{KNN-Neighborhood-Construction} is computationally simpler, \textsc{RNG-Neighborhood-Construction} typically produces graphs with better search performance at the expense of more complex neighborhood formation.
+In this paper, we performed our experiments with \textsc{RNG-Neighborhood-Construction}.
 
@@
 It iterates through each layer level and applies a chosen layer-merge algorithm (\textsc{NGM}, \textsc{IGTM}, or \textsc{CGTM}) to combine the corresponding layers $G^a_i$ and $G^b_i$.
-The layer-merge algorithms require access to the full HNSW structures $\mathcal{H}_a, \mathcal{H}_b$ to perform efficient searches using \textsc{HNSW-Search}.
+The layer-merge algorithms require access to the full HNSW structures $\mathcal{H}_a, \mathcal{H}_b$ to perform searches using \textsc{HNSW-Search}.
 
@@
 
-The process is then repeated for all vertices of graph $G^b_\ell$ (lines 9-12), searching for their potential neighbors in $G^a_\ell$.
+The same process is then repeated for all vertices of graph $G^b_\ell$ (lines 9-12).
 
@@
 \subsection{Intra Graph Traversal Merge}
-The most effort of \textsc{NGM} algorithm lies in the set of neighborhood candidates from the opposite graph utilising \textsc{HNSW-Search} procedure, that traverses layers graph from the top-level down to layer number $\ell$ every time.
-The number of computations can be reduced if we select the next vertex $v^*$ close to the previous one (line 15), instead of choosing it randomly.
-Thus, for new $v^*$ the neighborhood candidates will be  also close to the previous candidates set.
-To search this new neighborhood candidates we can use \textsc{LocalSearch} procedure, that traversing the same graph staring from the previous neighborhood candidates set $\mathcal{P}^b$ (line 11,14).
-In the line 14 in set $\mathcal{P}^b$ we keep only $M$-closest to $v^*$ candidates.\\
-In the line 15 almost at each iteration we try select new $v^*$ vertex close to the previous  $v^*$ that was not already processed.  
-To ensure that new $v^*$ vertex are not very far from the previous $v^*$ we bound the size of results of \textsc{LocalSearch} procedure controlling by "next\_step\_k" parameter.
-Once \textsc{LocalSearch} can't find enough close not processed vertex, in line 7 we choice new $v^*$ from the set $\mathcal{V}_{not\_done}$ randomly.
+The most effort of the \textsc{NGM} algorithm lies in obtaining the set of neighborhood candidates from the other graph utilising \textsc{HNSW-Search} procedure, that every time traverses layer graphs from the top-level down to layer number $\ell$.
+The number of computations can be reduced if we select the next vertex to process $v^*$ close to the previous one (line 15), instead of randomly choosing it.
+Thus, for the new $v^*$ the neighborhood candidates will also be close to the previous candidates set.
+To search these new neighborhood candidates we can use \textsc{LocalSearch} procedure, that traverses the same graph staring from the previous neighborhood candidates set $\mathcal{P}^b$ (lines 11).
+In line 14 in set $\mathcal{P}^b$ we keep only $M$-closest to $v^*$ candidates.\\
+In lines 15,16 we select new processing vertex $v^*$  close to the previous processing vertex $v^*$ that was not already processed.  
+To ensure that the new processing vertex $v^*$ is not very far from the previous $v^*$ we bound the size of the results of the \textsc{LocalSearch} procedure controlling by next\_step\_k parameter.
+Once \textsc{LocalSearch} can't find enough close not processed vertex (if condition in line 17), in line 7 we choice new $v^*$ from the set $\mathcal{V}_{not\_done}$ randomly.
 \\
-After we have processed all vertices from the the graph $G^a$. We do the same for vertices of graph $G^b$ (lines 21-22).
+After we have processed all vertices from the graph $G^a$. We do the same for the vertices of graph $G^b$ (lines 21-22).
 
@@
 
-Cross Graph Traversal Merge (\textsc{CGTM}) algorithm is similar to \textsc{IGTM} utilise \textsc{LocalSearch} procedure to reduce computation effort.
-The difference is that \textsc{IGTM} algorithm the next $v^*$ vertex choice from the same graph, while \textsc{CGTM} looks for new vertex $v^*$ in the both graphs $G^a$, and $G^b$.
-Thus in line 24  $v^*$  is chosen from the set $\mathcal{C}_{not\_done}$, which of not processed vertex from the both graph (line 21).
-The intuition laying behind \textsc{CGTM} algorithm is that allowing to choice new $v^*$ from the both graphs, we reduce the number of times what $v^*$ is chosen randomly, thereby minimise the number of using more expensive search procedure \textsc{HNSW-Search}.
+Cross Graph Traversal Merge (\textsc{CGTM}) algorithm is similar to \textsc{IGTM} utilizes \textsc{LocalSearch} procedure to reduce computation efforts.
+The difference is that \textsc{IGTM} algorithm chooses the next processing vertex $v^*$ from the same graph, while \textsc{CGTM} looks for the new processing vertex $v^*$ in both graphs $G^a$, and $G^b$.
+Thus in line 24 $v^*$ is chosen from the set $\mathcal{C}_{not\_done}$, which is one of the not processed vertex of both graph (line 21).
+The intuition underlying \textsc{CGTM} algorithm is that when we choose a new processing vertex $v^*$ from both graphs, we reduce the number of times when $v^*$ is chosen randomly, thus minimizing the number of times that we use the more expensive search procedure \textsc{HNSW-Search}.
 
@@
 On those subsets we built two HNSW indices $\mathcal{H}_a$ and $\mathcal{H}_b$ using the following parameters: $M = 16$, $M0$ = 32, ef\_construction = 32.
-These indices are then merged using simple insertion stategy textsc{SIGM} and the three proposed algorithms: textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
-We keep the same values of parameters $m = 16$, and $m0 = 32$ as the target neighborhood size in the merged graph.
+These indices are then merged using simple insertion strategy \textsc{SIGM} and the three proposed algorithms: \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
+We keep the same values of the parameters $m = 16$, and $m0 = 32$ as the target neighborhood size in the merged graph.
 
-In order to be independent of the implementation details and low level optimisation, we have done comparison of the algorithms based on the number of distance computation required for the merge process.
+In order to be independent of implementation details and low-level optimisations, we made a comparison of the algorithms based on the number of distance computations required for the merge process.
 
-Other important characteristics of the merging process is how accurate the merge is done.
+Other important characteristics of the merge process is how accurate the merge is done.
 In this case, the word ``merging accuracy'' can mean different things.
@@
 
-In our experiment as recall@5 we report an averaged value over all sequence of searches.
+In our experiment as recall@5 we report an averaged value over all sequences of searches.
 
@@
 
-The searching quality of the merged graph can be seen from the Fig.
+The searching quality of the merged graph represented in Fig.
 ~\ref{fig:search}.
@@
 
-Another important fact which ~\ref{fig:search} shows that the number of distance computations performed by the searching algorithm (Algorithm ~\ref{alg:hnsw_search}) for equal parameter $L$ are almost the same.
+Another important fact which ~\ref{fig:search} shows that the number of distance computations performed by the search algorithm (Algorithm ~\ref{alg:hnsw_search}) for equal parameter $L$ are almost the same.
 Therefore, the graph is better if it provides better search recall for a fixed value of the parameter $L$.
@@
 Finally, future work can be applied to develop a fast graph construction procedure using one of the merge algorithms.
-It seems very naturally to start from the set of small graphs and merge them in a recursive manner.
+It seems very natural to start from the set of small graphs and merge them in a recursive manner.
 

### 2025-05-06 paper/main.tex
improve references

@@
 \title{Three Algorithms for Merging Hierarchical Navigable Small World Graphs}
-\author[1]{Alexander Ponomarenko}
+\author[1,2]{Alexander Ponomarenko}
 \date{}
 
-% \affil[2]{YugaByte LLC}
 \affil[1]{HSE Univerity}
+\affil[2]{YugaByte LLC}
 
-\usepackage{natbib}
+\usepackage[backend=biber,style=numeric-comp]{biblatex}
+\addbibresource{references.bib}
+\usepackage{hyperref}
+
 \usepackage{graphicx}
@@
 
-% В итнтро добавть о том, что такое навигационные графы. Что такое hnsw. Какие операции он поддерживает. 
-% Раскрыть глобальную идею мерджа, что merge можно понимать как итеративный процесс из двух уопераций;
-% 1. Выбор вершины (мн) для которой мы намериваемся строить окрестность и для какого графа.
-% 2. Выбор кандидатов из каждого графа
-% 3. Построение окрестности (согласно одной из выбранных стратегий).
-% 4. Принятие решение о том, какую информацию передать на следущую итерацию
-
-% TODO для экспериментов. 
-% 2. Провести эксперименты для разных размерностей и размеров
-% 3. Следить за тем сколько операций "полного поиска было сделано" и сколько операций "local search".
-% 5. Сделать все теже эксперементы, но используя в качестве NeighborhoodConstruction k-nn
-
-\section{Introduction}
-
 %TODO здесь надо сказать, именно проважность задачи k-nn для прирожений.
@@
 В частности это может быть важно при compaction.
-Navigable small-world graphs have emerged as a cornerstone for efficient approximate nearest neighbor (ANN) search in high-dimensional spaces.
-This problem is crucial in fields such as recommendation systems, image retrieval, and natural language processing.
-Algorithms like Hierarchical Navigable Small World (HNSW) \cite{hnsw} and its predecessor Navigable Small World (NSW) \cite{nsw2011,nsw2012,nsw2014} provide robust and scalable frameworks for organizing data into graph structures that support fast similarity searches.
+
+\section{Introduction}
+\label{sec:intro}
+The k-nearest neighbor (k-NN) search problem has emerged as a fundamental computational challenge with critical applications across numerous domains.
+In today's data-driven landscape, efficient similarity search is essential for recommendation systems, computer vision, multimedia retrieval, and natural language processing applications.
+Recent advancements in generative AI, particularly Retrieval-Augmented Generation (RAG) systems, have further highlighted the importance of fast and accurate nearest neighbor search.
+RAG frameworks depend on efficient retrieval of relevant document fragments from large vector databases to ground language model outputs in factual information, making high-performance approximate nearest neighbor (ANN) search algorithms increasingly vital.
+
+%As data volumes continue to grow exponentially, traditional exact k-NN methods struggle with the computational demands of high-dimensional spaces.
+This has driven significant research interest in approximate methods that sacrifice perfect accuracy for substantial improvements in query speed and memory efficiency.
+Among these approaches, navigable small-world graphs have emerged as a particularly effective framework for organizing data into structures that support fast similarity searches.
 
 \subsection{Nearest Neighbor Search Problem}
-
-% Formally, the nearest neighbor search problem can be defined as follows: Given a set of points $P = \{p_1, p_2, ..., p_n\}$ in a metric space $(X, \rho)$, where $\rho$ is a distance function, and a query point $q \in X$, find the point $p^* \in P$ such that:
-
-% $$p^* = \underset{p \in P}{\mathrm{argmin}} \rho(q, p)$$
 
@@
 
-In high-dimensional spaces, this problem becomes computationally challenging due to the ``curse of dimensionality.'' Exact solutions often degrade to linear scan as dimensionality increases, making them impractical for large-scale applications.
+%In high-dimensional spaces, this problem becomes computationally challenging due to the ``curse of dimensionality.'' Exact solutions often degrade to linear scan as dimensionality increases, making them impractical for large-scale applications.
 This has led to the development of approximate $k$-nearest neighbor (ANN) search methods, which trade perfect accuracy for substantial improvements in computational efficiency.
@@
     \item \textbf{Space-partitioning trees}: These methods divide the vector space into hierarchical regions, enabling logarithmic search complexity in low dimensions.
-Notable examples include KD-trees \cite{Bentley1990k}, R-trees \cite{guttman1984r}, VP-trees \cite{yianilos1993}, M-trees \cite{ciaccia1997m}, and Cover trees \cite{beygelzimer2006cover}.
+Notable examples include KD-trees \cite{bentley1990k}, R-trees \cite{guttman1984r}, VP-trees \cite{yianilos1993}, M-trees \cite{ciaccia1997m}, and Cover trees \cite{beygelzimer2006cover}.
 A comprehensive overview of these approaches for metric spaces is provided by Zezula et al.
@@
 
-HNSW \cite{hnsw} extends the navigable small-world concept by organizing points into a hierarchy of navigable graphs also named as layers.
-Each layer contains a progressively smaller subset of the data points.
+Hierarchical Navigable Small World (HNSW) \cite{hnsw} is a state of the art navigable graph-based method.
+It demonstrates superior performance on modern ANN benchmarks \cite{aumuller2020ann}.
+Due to relative simplicity, and efficiency, the HNSW algorithm has become widely used.
+It has been implemented in popular libraries for ANN-Search \cite{faiss}, \cite{nmspacelib}, and it is implemented as vector index in many database systems including PostgreSQL (vector extension), ORACLE \cite{OracleHNSW2023}, and other vector search oriented databases such as Milvus \cite{MilvusHNSW2024}, Zilliz \cite{ZillizHNSW2022}, Weaviate \cite{WeaviateVectorIndex2024}.
+
+%Among graph-based methods, Hierarchical Navigable Small World (HNSW)  and its predecessor Navigable Small World (NSW) \cite{nsw2011,nsw2012,nsw2014} have demonstrated superior performance in modern ANN benchmarks.
+These approaches leverage the small-world phenomenon to create graph structures with both short-range and long-range connections, enabling efficient navigation through high-dimensional data spaces.
+
+HNSW \cite{hnsw} extends the navigable graphs concept by organizing points into a hierarchy of navigable graphs also named as layers.
+Each layer contains a progressively smaller subset of data points.
 The bottommost layer contains all points, while higher layers form a sparse representation of the data space.
-Graphs of all levels glued together form a navigable small world graph like NSW \cite{nsw2011}.
-An explicit separation small world graph to layers allows to reduce number of computation in search while traversing vertexes with high degrees also called hubs.
-
-HNSW supports two primary operations: \textbf{search} (see section \ref{subasec: hnsw-search}) and  \textbf{insertion}.
-In the Fig.
-~\ref{fig:hnsw} is a schematic example of multilayered structure of HNSW.
+Graphs of all levels glued together form a navigable small world graph similar to NSW  \cite{nsw2011,nsw2012,nsw2014}.
+An explicit separation set of edges of different lengths to layers helps to reduce the number of computations in search while traversing vertexes with high degrees, also called hubs.
+A schematic example of HNSW structure presented in Fig.
+\ref{fig:hnsw}.
 
@@
 Each layer is a navigable graph that suit for ``greedy-like'' traversal algorithms.
-Graphs of all levels glued together form a navigable small world graph like similar to NSW.
-An explicit separation small world graph to layers allows to reduce number of computation in search while traversing vertexes with high degrees also called hubs.}
+Graphs of all levels glued together form a navigable small world graph similar to NSW.
+An explicit separation set of edges of different lengths to layers helps to reduce the number of computations in search while traversing vertexes with high degrees, also called hubs.
+}
 \label{fig:hnsw}
@@
 
-\subsection{Graph Merging as an Iterative Process}
-One of the key challenges in maintaining these graph-based indices is efficiently combining data from multiple sources.
-While HNSW supports dynamic insertions, merging entire graphs as coherent structures presents a different challenge.
-The quality of the merged graph directly impacts search performance, and naive approaches may lead to suboptimal neighborhood connections or excessive computational costs during the merge operation.
+\subsection{Graph Merging Challenge}
 
-It is worth noting that graph merging can be used as an analog to compaction operations in database systems.
+While HNSW supports efficient dynamic insertions and deletions, an important operational challenge arises in practical systems: merging multiple independently constructed graph indices.
+This problem is particularly relevant in distributed systems, incremental indexing scenarios, and database compaction operations.
 In graph-based indices, deletion operations are often implemented by simply marking vertices as deleted without actually removing them, to avoid disrupting the graph's connectivity.
 This approach prevents the computationally expensive task of rebuilding neighborhoods for affected vertices.
-Moreover, a true deletion might trigger cascading neighborhood reconstructions that could potentially affect the entire graph, which is unacceptable from a performance perspective.
-Graph merging provides a way to periodically compact the structure by creating a fresh graph that excludes deleted vertices.
+However, as deletions accumulate, the index becomes increasingly sparse and less efficient.
+
+Graph merging provides a way to periodically compact the structure by creating a fresh graph that excludes deleted vertices while preserving the navigability properties essential for efficient search.
+Additionally, in distributed computing environments, merging separately constructed indices becomes necessary for building unified search structures over partitioned datasets.
+
+\subsection{Graph Merging as an Iterative Process}
+%One of the key challenges in merging graph-based indices is efficiently combining data from multiple sources while preserving the navigable small-world properties.
+The quality of the merged graph directly impacts search performance, and naive approaches may lead to suboptimal neighborhood connections or excessive computational costs during the merge operation.
 
@@
 
-In this work, we propose three novel algorithms based on this iterative framework—\textbf{Naive Graph Merge (NGM)}, \textbf{Intra Graph Traversal Merge (IGTM)}, and \textbf{Cross Graph Traversal Merge (CGTM)}—for efficiently merging navigable graphs.
+In this work, we propose three novel algorithms based on this iterative framework—\textbf{Naive Graph Merge (NGM)}, \textbf{Intra Graph Traversal Merge (IGTM)}, and \textbf{Cross Graph Traversal Merge (CGTM)}—for efficiently merging HNSW graphs.
 These algorithms differ primarily in how they implement each step of the iterative process.
-For instance, NGM and IGTM select the next vertex for neighborhood construction only from the same graph as the current vertex, promoting locality within original graphs.
-In contrast, CGTM implements the idea that the next vertex for neighborhood construction can be selected from any input graph, enabling more thorough exploration of the combined space.
-While our methods primarily target HNSW, they are generalizable to other graph-based data structures such as NSW \cite{nsw2011} and NSG \cite{NSG}.
-These algorithms preserve graph navigability and maintain the structural integrity necessary for fast ANN search while offering different trade-offs between merge accuracy and computational efficiency.
+For instance, NGM and IGTM select the next vertex for neighborhood construction only from the same graph as the current vertex.
+In contrast, CGTM implements the idea that the next vertex for neighborhood construction can be selected from any input graph.
 
-The rest of this paper is organized as follows: Section 2 describes the search algorithms for navigable graphs that we use in merge procedures.
-Section 3 presents neighborhood construction strategies that are essential for both graph building and merging.
-Section 4 introduces our three graph merging algorithms in detail.
-Section 5 presents experimental results comparing the proposed methods.
-Finally, Section 6 concludes the paper and discusses directions for future research.
+The rest of this paper is organized as follows.
+Section 2 describes the search algorithms for HNSW graphs that we use in merge procedures.
+Section \ref{sec:neighborhood} recalls neighborhood construction strategies that are essential for both graph building and merging.
+Section \ref{sec:merge} introduces our three graph merging algorithms in detail.
+Section \ref{sec:experimens} presents computational experimental results.
+We give references to similar works in Section \ref{sec:relatedworks}.
+Finally, Section \ref{sec:conclusion} concludes the paper and discusses directions for future research.
 
 \section{Search}
+\label{sec:search}
 
@@
 Below, we describe two key search algorithms: \textsc{LocalSearch} for exploring a single graph layer, and \textsc{HNSW-Search} for hierarchical multi-layer traversal.
-
-%
 
@@
 \section{Neighborhood Construction Strategies}
+\label{sec:neighborhood}
 
@@
 \section{Merge Algorithms}
+\label{sec:merge}
 
@@
 \section{Computational Experiments}
+\label{sec:experimens}
 
@@
 
+\section{Related Works on Graph Merging}
+\label{sec:relatedworks}
+While working on the paper, we have found that just recently researchers from \textbf{Elastic Search Lab} published a blog post \cite{ElasticHNSW2023} where they proposed fast merge algorithm, which also utilizes information about closeness of object in all graphs.
+The authors presented a scheme where a subset of vertices $J$ ``join set'' from the smaller graph is selected by greedy heuristic and inserted into the larger graph using standard HNSW insertion procedure.
+For the remaining vertices from the small graph, their already inserted neighbors and those neighbors' connections in the large graph are used to perform a limited beam search (named FAST-SEARCH-LAYER) to find connection candidates.
+This approach limits full insertions and accelerates the merging process.
+
+Ideologically approach of \cite{ElasticHNSW2023} is similar to our work, because the authors utilize information about closeness of object in all graphs, and use light-weighted version of the search algorithm FAST-SEARCH-LAYER (in our case it is \textsc{LOCAL SEARCH}.
+The ``join set'' $J$ from some point of view is similar to the set vertices for which \textsc{IGTM} and \textsc{CGTM} run the standard HNSW search.
+However, in \cite{ElasticHNSW2023} the ``join set'' $J$ is determined separately at the first stage by a greedy heuristic that tries to find a good cover set, while \textsc{IGTM} and \textsc{CGTM} establish the vertices for which to run \textsc{HNSW-Search} during the work in the main cycle in a more homogeneous way.
+
+The other work which is relatively close to the present topic is a paper by Zhao and co-authors \cite{knnmerge}.
+They proposed two algorithms: Symmetric Merge (S-Merge) for combining two k-NN graphs and Joint Merge (J-Merge) for incrementally extending an existing graph with new points.
+These methods effectively combine graphs built on different subsamples while maintaining high search accuracy and accelerating computations.
+S-Merge discards half the neighbors in each list, while J-Merge leverages the hierarchical structure of the graph.
+
+The performance comparison of the different merge algorithms is a subject of future research.
+
 \section{Conclusion and Future Work}
+\label{sec:conclusion}
 
@@
 
-Our experimental results demonstrate that both \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to the naive approach while maintaining comparable search accuracy.
+Our experimental results demonstrate that \textsc{IGTM} and \textsc{CGTM} significantly reduce the number of distance computations compared to the naive approach while maintaining comparable search accuracy.
 Our evaluation on the dataset of 1 million 128-dimensional vectors (SIFT1M) shows that \textsc{IGTM} and \textsc{CGTM} are more than 3 times faster than the straightforward insertion strategy \textsc{SIGM}, with minimal impact on recall performance.
@@
 We had anticipated that \textsc{CGTM} would be more efficient since it can select the next vertex for neighborhood construction from both graphs, theoretically reducing the chances of getting stuck and thus requiring fewer costly standard search operations.
-However, it appears that the overhead of selecting the next processing vertex, for which a neighborhood is forming, in \textsc{CGTM} is too computationally expensive, and these costs are not offset by the reduction in searches when ``getting stuck'' with selecting the next close vertex to process.
+However, it appears that the overhead of selecting the next processing vertex, for which a neighborhood is forming, in \textsc{CGTM} is too computationally expensive, and these costs are not offset by the reduction in searches when ``gets stuck'' with selecting the next close vertex to process.
 The study of this fact is a subject for further work.
@@
 
-\bibliographystyle{plain}
-\bibliography{references}
+\printbibliography
 \end{document}

### 2025-05-06 paper/main.tex
add abstract to the paper

@@
 
-%TODO здесь надо сказать, именно проважность задачи k-nn для прирожений.
-Упомянть про RAG, а далее уже перейти к формальной постановке задаче и какие методы для её решения бывают.
-Далее надо сказать про после того, как было упомянуто про графовые методы, сказать что там етсь задача объединения графов построенных на разных данных.
-В частности это может быть важно при compaction.
+\begin{abstract}
+This paper addresses the challenge of merging hierarchical navigable small world (HNSW) graphs, a critical operation for distributed systems, incremental indexing, and database compaction.
+We propose three algorithms for this task: Naive Graph Merge (NGM), Intra Graph Traversal Merge (IGTM), and Cross Graph Traversal Merge (CGTM).
+These algorithms differ in their approach to vertex selection and candidate collection during the merge process.
+We conceptualize graph merging as an iterative process with four key steps: processing vertex selection, candidate collection, neighborhood construction, and information propagation.
+Our experimental evaluation on the SIFT1M dataset demonstrates that IGTM and CGTM significantly reduce computational costs compared to naive approaches, requiring up to 70\% fewer distance computations while maintaining comparable search accuracy.
+Surprisingly, IGTM outperforms CGTM in efficiency, contrary to our initial expectations.
+The proposed algorithms enable efficient consolidation of separately constructed indices, supporting critical operations in modern vector databases and retrieval systems that rely on HNSW for similarity search.
+\end{abstract}
+
+\textbf{Keywords:} approximate nearest neighbor search, hierarchical navigable small world, graph merging, vector databases, information retrieval
 

### 2025-05-21 paper/main.tex
fix affiliations
Apply Sandeep comments to the paper

@@
 
-\affil[1]{HSE Univerity}
-\affil[2]{YugaByte LLC}
+\affil[1]{Laboratory LATNA, HSE University}
+\affil[2]{Yugabyte, Inc}
 
@@
     \item \textbf{Mapping-based methods}: These techniques transform the original data into representations that allow for efficient search.
-This category includes Locality-Sensitive Hashing (LSH) \cite{Indyk1998,Gionis1999,Datar2004,Andoni2006,Andoni2015} and Product Quantization approaches \cite{Jegou2011,Ge2013,Norouzi2013}, which encode vectors into compact codes that approximate distances between points.
+This category includes Locality-Sensitive Hashing (LSH) \cite{Indyk1998,Gionis1999,Datar2004,Andoni2006,Andoni2015} and Product Quantization approaches \cite{Jegou2011,Ge2013,Norouzi2013}, which encode vectors into compact codes that approximate the distances between points.
 
     \item \textbf{Navigable graph-based methods}: The most recent and currently state-of-the-art approach that constructs graph structures where vertices represent data points and edges connect similar points.
-These graphs are designed to allow ``greedy-like'' algorithms to perform directed traversal through the data space.
+These graphs are designed to allow ``greedy-like'' algorithms to perform a directed traversal through the data space.
 \end{enumerate}
@@
 
-HNSW \cite{hnsw} extends the navigable graphs concept by organizing points into a hierarchy of navigable graphs also named as layers.
+HNSW \cite{hnsw} extends the navigable graphs concept by organizing points in a hierarchy of navigable graphs also named layers.
 Each layer contains a progressively smaller subset of data points.
-The bottommost layer contains all points, while higher layers form a sparse representation of the data space.
+The bottommost layer contains all points, while the upper layers form a sparse representation of the data space.
 Graphs of all levels glued together form a navigable small world graph similar to NSW  \cite{nsw2011,nsw2012,nsw2014}.
@@
 \begin{enumerate}
-    \item \textbf{Processing Vertex Selection}: Choosing a vertex $v^*$ for which we aim to construct a neighborhood in the merged graph, and determining which original graph it belongs to.
+    \item \textbf{Processing Vertex Selection}: Choosing a vertex $v^*$ for which we aim to construct a neighborhood in the merged graph and determining which original graph it belongs to.
 
@@
 We start from description of a general algorithm for merging multi layers structure.
-We now present three algorithms for merging HNSW graphs: \textsc{SIGM},  \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
+We now present three algorithms for merging HNSW graphs: \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
 These algorithms operate layer by layer.
@@
 \subsection{Simple Insertion Graph Merge}
-Before we start, we need provide a short descriptopn of simple insertion graph merge (SIGM).
+Before we start, we need to provide a short description of simple insertion graph merge (SIGM).
 It is a default merging strategy that uses an insertion operation to merge graphs.
 It takes the largest graph and sequentially inserts data from other graph.
-Actually, it is not merging procedure, but we use it as a basic benchmark.
+Actually, it is not a merging procedure, but we use it as a basic benchmark.
 
@@
 
-Note that the merging of different layers (iterations of the loop in Algorithm~\ref{alg:general_merge}) is independent and can potentially be parallelized.
+% Note that the merging of different layers (iterations of the loop in Algorithm~\ref{alg:general_merge}) is independent and can potentially be parallelized.
 
@@
 \subsection{Intra Graph Traversal Merge}
-The most effort of the \textsc{NGM} algorithm lies in obtaining the set of neighborhood candidates from the other graph utilising \textsc{HNSW-Search} procedure, that every time traverses layer graphs from the top-level down to layer number $\ell$.
-The number of computations can be reduced if we select the next vertex to process $v^*$ close to the previous one (line 15), instead of randomly choosing it.
-Thus, for the new $v^*$ the neighborhood candidates will also be close to the previous candidates set.
-To search these new neighborhood candidates we can use \textsc{LocalSearch} procedure, that traverses the same graph staring from the previous neighborhood candidates set $\mathcal{P}^b$ (lines 11).
-In line 14 in set $\mathcal{P}^b$ we keep only $M$-closest to $v^*$ candidates.\\
-In lines 15,16 we select new processing vertex $v^*$  close to the previous processing vertex $v^*$ that was not already processed.  
-To ensure that the new processing vertex $v^*$ is not very far from the previous $v^*$ we bound the size of the results of the \textsc{LocalSearch} procedure controlling by next\_step\_k parameter.
-Once \textsc{LocalSearch} can't find enough close not processed vertex (if condition in line 17), in line 7 we choice new $v^*$ from the set $\mathcal{V}_{not\_done}$ randomly.
-\\
-After we have processed all vertices from the graph $G^a$. We do the same for the vertices of graph $G^b$ (lines 21-22).
 
@@
 
+The most effort of the \textsc{NGM} algorithm lies in obtaining the set of neighborhood candidates from the other graph utilizing the \textsc{HNSW-Search} procedure, which every time traverses the layer graphs from the top level down to the layer number $\ell$.
+The number of computations can be reduced if we select the next vertex to process $v^*$ close to the previous one (line 15 of the algorithm ~\ref{alg:IGTM}), instead of randomly choosing it.
+Thus, for the new $v^*$ the neighborhood candidates will also be close to the previous candidates set.
+To search for these new neighborhood candidates we can use the \textsc{LocalSearch} procedure which traverses the same graph staring from the previous neighborhood candidates set $\mathcal{P}^b$ (lines 11).
+In line 14 in the set $\mathcal{P}^b$ we keep only $M$-closest to $v^*$ candidates.\\
+In lines 15,16, we select a new processing vertex $v^*$  close to the previous processing vertex $v^*$ that was not already processed.  
+To ensure that the new processing vertex $v^*$ is not very far from the previous $v^*$ we bound the size of the results of the \textsc{LocalSearch} procedure controlling by next\_step\_k parameter.
+Once \textsc{LocalSearch} cannot find enough close unprocessed vertex (if condition in line 17), in line 7 we choose a new $v^*$ from the set $\mathcal{V}_{not\_done}$ randomly.\\
+After we have processed all vertices from the graph $G^a$. We do the same for the vertices of graph $G^b$ (lines 21-22).
+
 \subsection{Cross Graph Traversal Merge}
 
-Cross Graph Traversal Merge (\textsc{CGTM}) algorithm is similar to \textsc{IGTM} utilizes \textsc{LocalSearch} procedure to reduce computation efforts.
-The difference is that \textsc{IGTM} algorithm chooses the next processing vertex $v^*$ from the same graph, while \textsc{CGTM} looks for the new processing vertex $v^*$ in both graphs $G^a$, and $G^b$.
+Cross Graph Traversal Merge (\textsc{CGTM}) algorithm is similar to \textsc{IGTM} utilizes the \textsc{LocalSearch} procedure to reduce computation efforts.
+The difference is that the \textsc{IGTM} algorithm chooses the next processing vertex $v^*$ from the same graph, while \textsc{CGTM} looks for the new processing vertex $v^*$ in both graphs $G^a$, and $G^b$.
 Thus in line 24 $v^*$ is chosen from the set $\mathcal{C}_{not\_done}$, which is one of the not processed vertex of both graph (line 21).
@@
 We divided it into two disjoint subsets of 500k vectors.
-On those subsets we built two HNSW indices $\mathcal{H}_a$ and $\mathcal{H}_b$ using the following parameters: $M = 16$, $M0$ = 32, ef\_construction = 32.
-These indices are then merged using simple insertion strategy \textsc{SIGM} and the three proposed algorithms: \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
-We keep the same values of the parameters $m = 16$, and $m0 = 32$ as the target neighborhood size in the merged graph.
+In these subsets, we built two HNSW indices $\mathcal{H}_a$ and $\mathcal{H}_b$ using the following parameters: $M = 16$, $M0$ = 32, ef\_construction = 32.
+Then these indices were merged using a simple insertion strategy \textsc{SIGM} and the three proposed algorithms: \textsc{NGM}, \textsc{IGTM}, and \textsc{CGTM}.
+We keep the same values of the parameters $m = 16$ and $m0 = 32$ as the target neighborhood size in the merged graph.
 
@@
 
-Formally, for a given query point \(q\), let \(P_k(q)\) denote the set of \(k\) true nearest neighbors, and let \(A_k(q)\) be the set of top \(k\) points returned by the algorithm.
-Then, the recall@k is defined as:
+Formally, for a given query point \(q\), let \(P_k(q)\) denote the set of \(k\) true nearest neighbors, and let \(A_k(q)\) be the set of the top \(k\) points returned by the algorithm.
+Then, recall@k is defined as:
 
@@
 
-The searching quality of the merged graph represented in Fig.
+The searching quality of the merged graph is represented in Fig.
 ~\ref{fig:search}.
@@
 \begin{itemize}
-    \item\textsc{IGTM} and \textsc{IGTM}  in our experiment setup  achieve recall better than \textsc{SIGM} with parameter ef\_construciton = 24;
+    \item\textsc{IGTM} and \textsc{CGTM} in our experiment setup  achieve recall better than \textsc{SIGM} with parameter ef\_construciton = 24;
     \item \textsc{NGM} and \textsc{SIGM} perform the most computations, but achieves the highest recall, as it exhaustively reconstructs each neighborhood;
@@
 \label{sec:relatedworks}
-While working on the paper, we have found that just recently researchers from \textbf{Elastic Search Lab} published a blog post \cite{ElasticHNSW2023} where they proposed fast merge algorithm, which also utilizes information about closeness of object in all graphs.
+While working on the paper, we have found that just recently researchers from \textbf{Elastic Search Lab} published a blog post \cite{ElasticHNSW2023} where they proposed a fast merge algorithm, which also utilizes information about closeness of object in all graphs.
 The authors presented a scheme where a subset of vertices $J$ ``join set'' from the smaller graph is selected by greedy heuristic and inserted into the larger graph using standard HNSW insertion procedure.

### 2025-05-21 paper/main.tex
Add link to repo into the paper

@@
 \]
+In our experiment as recall@5 we report an averaged value over all sequences of searches.
 
-In our experiment as recall@5 we report an averaged value over all sequences of searches.
+All source codes of our implementation can be found in a public GitHub repository \footnote{\url{https://github.com/aponom84/merging-navigable-graphs}}.
 
